---
title: "Open Finance (Brazil) — Quantitative Analysis (PIX & Adoption Focus)"
subtitle: "Analyzing PIX Transaction Growth, Customer Adoption, and Macro Determinants"
author: "Pranav Mehta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

```{r include=FALSE}
required_pkgs <- c(
  "tidyverse", "readxl", "lubridate",
  "scales", "broom", "sandwich", "lmtest", "zoo"
)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(broom)
library(sandwich)
library(lmtest)
library(zoo)
```

# Objective

This notebook analyzes the growth dynamics of Brazil's Open Finance ecosystem using two key proxies:

1. **PIX transaction volumes** — Brazil's instant payment system, launched in November 2020, which has become the backbone of digital financial infrastructure.
2. **Open Finance customer counts** — the number of customers with active data-sharing consents across participating institutions.

We examine how these adoption metrics relate to macroeconomic conditions, specifically:

- **Selic** (the Central Bank of Brazil's policy interest rate)
- **IBC-Br** (the Central Bank's monthly economic activity index, a GDP proxy)

# Background

## PIX

PIX is Brazil's instant payment system, launched by the Central Bank in November 2020. It enables real-time, 24/7 transfers between individuals, businesses, and government at zero cost to end users. PIX adoption has been explosive — from zero to billions of monthly transactions within a few years — making it one of the most successful instant payment deployments globally.

In the context of Open Finance, PIX serves as critical infrastructure: it enables payment initiation via third-party providers, a core pillar of the Open Finance framework.

## Selic

Selic is Brazil's policy interest rate target. It is the primary instrument used by the Central Bank of Brazil (BCB) to influence inflation and economic activity:

- When Selic rises, borrowing costs increase and economic activity often slows.
- When Selic falls, financial conditions loosen and activity may accelerate.

## IBC-Br

The IBC-Br (Índice de Atividade Econômica do Banco Central) is a monthly economic activity index published by the BCB. It serves as a high-frequency proxy for GDP, capturing the trajectory of economic output between quarterly GDP releases.

# Data Sources

All data come from a single Excel workbook (`data/data_monthly.xlsx`) containing two sheets:

**Sheet 1** — PIX, Customers, Selic, IBC-Br (January 2023 – January 2025)

- `month`: calendar month
- `pix`: PIX transaction count
- `cust`: Open Finance customer count
- `selic`: Selic rate (% p.a.)
- `ibc_br`: IBC-Br index

**Sheet 2** — PIX, Selic, IBC-Br (November 2020 – December 2024)

- `month`: calendar month
- `pix`: PIX transaction count
- `selic`: Selic rate (% p.a.)
- `ibc_br`: IBC-Br index

Sheet 2 provides the longer time series needed for year-over-year growth calculations, while Sheet 1 adds customer adoption data for a shorter but richer panel.

# Load and Process Data

```{r load-data}
# Read both sheets from the Excel workbook
df_long <- read_excel("data/data_monthly.xlsx", sheet = 2) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_short <- read_excel("data/data_monthly.xlsx", sheet = 1) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

cat("Long series (Sheet 2):", nrow(df_long), "months,",
    as.character(min(df_long$month)), "to", as.character(max(df_long$month)), "\n")
cat("Short series (Sheet 1):", nrow(df_short), "months,",
    as.character(min(df_short$month)), "to", as.character(max(df_short$month)), "\n")
```

```{r merge-data}
# Merge customer data from Sheet 1 into the longer series
df_merged <- df_long %>%
  left_join(df_short %>% select(month, cust), by = "month")

cat("Merged dataset:", nrow(df_merged), "months\n")
cat("Months with customer data:", sum(!is.na(df_merged$cust)), "\n")

summary(df_merged)
```

# Feature Engineering

We compute log transformations and year-over-year (YoY) growth rates using 12-month log differences:

$$
g^{yoy}_t \approx 100 \times \left(\ln X_t - \ln X_{t-12}\right)
$$

For PIX and IBC-Br, this captures the pace of expansion controlling for seasonality.

```{r features}
df_feat <- df_merged %>%
  mutate(
    ln_pix    = log(pix),
    ln_ibc_br = log(ibc_br),

    # YoY growth rates (require 12 months of history)
    g_yoy_pix    = 100 * (ln_pix - lag(ln_pix, 12)),
    g_yoy_ibc_br = 100 * (ln_ibc_br - lag(ln_ibc_br, 12)),

    # Month-over-month growth (for shorter-horizon analysis)
    g_mom_pix    = 100 * (ln_pix - lag(ln_pix, 1)),
    g_mom_ibc_br = 100 * (ln_ibc_br - lag(ln_ibc_br, 1))
  )

# Add customer features where available
df_feat <- df_feat %>%
  mutate(
    ln_cust      = ifelse(!is.na(cust), log(cust), NA_real_),
    g_yoy_cust   = 100 * (ln_cust - lag(ln_cust, 12)),
    g_mom_cust   = 100 * (ln_cust - lag(ln_cust, 1))
  )

cat("Observations with YoY PIX growth:", sum(!is.na(df_feat$g_yoy_pix)), "\n")
cat("Observations with YoY customer growth:", sum(!is.na(df_feat$g_yoy_cust)), "\n")
```

# Descriptive Analysis

## PIX Transaction Volumes

```{r plot-pix-levels}
df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Monthly Transaction Volume",
       subtitle = "Billions of transactions",
       x = NULL, y = "Transactions (billions)") +
  theme_minimal()
```

## PIX YoY Growth

```{r plot-pix-yoy}
df_feat %>%
  filter(!is.na(g_yoy_pix)) %>%
  ggplot(aes(month, g_yoy_pix)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "PIX Year-over-Year Transaction Growth",
       x = NULL, y = "YoY Growth (%)") +
  theme_minimal()
```

## Open Finance Customer Counts

```{r plot-cust}
df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "darkgreen", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Open Finance Customer Counts",
       subtitle = "Millions of customers",
       x = NULL, y = "Customers (millions)") +
  theme_minimal()
```

## Selic Rate

```{r plot-selic}
df_feat %>%
  ggplot(aes(month, selic)) +
  geom_line(color = "firebrick", linewidth = 1) +
  labs(title = "Selic Policy Rate",
       x = NULL, y = "Selic (% p.a.)") +
  theme_minimal()
```

## IBC-Br Economic Activity Index

```{r plot-ibc}
df_feat %>%
  ggplot(aes(month, ibc_br)) +
  geom_line(color = "darkorange", linewidth = 1) +
  labs(title = "IBC-Br Economic Activity Index",
       x = NULL, y = "IBC-Br") +
  theme_minimal()
```

## Combined Dashboard

```{r plot-dashboard, fig.height=10}
p1 <- df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Transactions (Bn)", x = NULL, y = NULL) +
  theme_minimal()

p2 <- df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "OF Customers (Mn)", x = NULL, y = NULL) +
  theme_minimal()

p3 <- df_feat %>%
  ggplot(aes(month, selic)) +
  geom_line(color = "firebrick", linewidth = 0.8) +
  labs(title = "Selic (%)", x = NULL, y = NULL) +
  theme_minimal()

p4 <- df_feat %>%
  ggplot(aes(month, ibc_br)) +
  geom_line(color = "darkorange", linewidth = 0.8) +
  labs(title = "IBC-Br", x = NULL, y = NULL) +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 1)
```

## Correlation Matrix

```{r correlation}
df_corr <- df_feat %>%
  select(pix, cust, selic, ibc_br) %>%
  drop_na()

cor_mat <- cor(df_corr)
round(cor_mat, 3)
```

# Regression Framework

We estimate time-series OLS regressions where the dependent variable is **PIX YoY transaction growth**, and regressors include:

- **Selic** (policy rate) — captures monetary conditions
- **IBC-Br YoY growth** — captures real economic activity

## Baseline Regression Equation

$$
g^{yoy}_{PIX,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\varepsilon_t
$$

where:

- $g^{yoy}_{PIX,t}$ is the YoY growth rate of PIX transactions at time $t$
- $Selic_t$ is the policy interest rate
- $g^{yoy}_{IBC\text{-}Br,t}$ is the YoY growth rate of the IBC-Br index
- $\varepsilon_t$ is the error term

## Interpretation

- $\beta_1$ captures the association between monetary conditions and PIX growth. A negative coefficient would suggest tighter monetary policy is associated with slower PIX adoption growth.
- $\beta_2$ captures **procyclicality**: whether PIX transaction growth accelerates during stronger economic conditions.

## Newey–West Standard Errors

Monthly time-series data commonly exhibit autocorrelation and heteroskedasticity. We use **Newey–West (HAC)** standard errors to produce consistent inference under these conditions. Given our sample size, we use a lag truncation parameter proportional to the data frequency.

# Baseline Model: PIX Growth ~ Selic + IBC-Br Growth

```{r model-data-baseline}
df_model_base <- df_feat %>%
  select(month, g_yoy_pix, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("Baseline sample:", nrow(df_model_base), "months\n")
cat("Sample range:", as.character(min(df_model_base$month)),
    "to", as.character(max(df_model_base$month)), "\n")
summary(df_model_base)
```

## Results: Baseline

```{r reg-baseline}
m_base <- lm(g_yoy_pix ~ selic + g_yoy_ibc_br, data = df_model_base)

# Newey-West HAC standard errors
nw_lag <- floor(0.75 * nrow(df_model_base)^(1/3))
cat("Newey-West lag truncation:", nw_lag, "\n")

nw_base <- NeweyWest(m_base, lag = nw_lag, prewhite = FALSE)

coeftest(m_base, vcov = nw_base)
```

```{r reg-baseline-summary}
cat("\nR-squared:", round(summary(m_base)$r.squared, 4), "\n")
cat("Adj. R-squared:", round(summary(m_base)$adj.r.squared, 4), "\n")
```

## Interpretation (Baseline)

Key considerations:

- **Selic**: A negative coefficient would indicate that higher interest rates are associated with slower PIX growth, consistent with tighter monetary conditions dampening digital payment adoption. However, PIX adoption may be driven primarily by structural and behavioral shifts rather than interest rate cycles.
- **IBC-Br growth**: A positive coefficient would suggest PIX growth is procyclical — transaction volumes grow faster when the economy is expanding.
- **R-squared**: Indicates how much of the variation in PIX growth is explained by macro variables alone. A low R-squared would suggest PIX growth is driven primarily by non-macro factors (adoption dynamics, network effects, regulatory mandates).

# Extended Model: Adding Open Finance Customer Growth

## Motivation

The baseline model captures macro determinants of PIX growth. We now add **Open Finance customer counts** (log-transformed) to test whether broader Open Finance adoption is associated with PIX transaction growth beyond what macro conditions explain.

Customer data is available from January 2023, so this model uses a shorter sample.

## Extended Regression Equation

$$
g^{yoy}_{PIX,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

where $\ln(\text{Customers}_t)$ is the natural log of Open Finance customer counts at time $t$.

- $\beta_3 > 0$: greater Open Finance adoption is associated with faster PIX growth, controlling for macro conditions.
- $\beta_3 \leq 0$: no detectable positive association.

## Extended Modeling Dataset

```{r model-data-extended}
df_model_ext <- df_feat %>%
  select(month, g_yoy_pix, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("Extended model sample:", nrow(df_model_ext), "months\n")
cat("Sample range:", as.character(min(df_model_ext$month)),
    "to", as.character(max(df_model_ext$month)), "\n")
summary(df_model_ext)
```

## Results: Extended Model

```{r reg-extended}
m_ext <- lm(g_yoy_pix ~ selic + g_yoy_ibc_br + ln_cust, data = df_model_ext)

nw_lag_ext <- max(1, floor(0.75 * nrow(df_model_ext)^(1/3)))
cat("Newey-West lag truncation:", nw_lag_ext, "\n")

nw_ext <- NeweyWest(m_ext, lag = nw_lag_ext, prewhite = FALSE)

coeftest(m_ext, vcov = nw_ext)
```

```{r reg-extended-summary}
cat("\nR-squared:", round(summary(m_ext)$r.squared, 4), "\n")
cat("Adj. R-squared:", round(summary(m_ext)$adj.r.squared, 4), "\n")
```

## Interpretation (Extended Model)

Key considerations:

1. **Sample restriction**: The extended model uses only the months with customer data (~12–13 months with YoY growth available), substantially fewer than the baseline. This reduces statistical power and increases standard errors.

2. **Collinearity with trend**: `ln_cust` is likely monotonically increasing, which means it is correlated with any time trend. A significant coefficient may reflect a general trend rather than a causal relationship.

3. **Causal interpretation**: This is an observational regression. A positive $\hat{\beta}_3$ does **not** imply that Open Finance adoption *caused* PIX growth. Plausible confounders include:
   - Reverse causality (PIX growth driving OF adoption)
   - Omitted variables (post-pandemic digitization driving both)
   - Regulatory mandates affecting both simultaneously

4. **Magnitude**: Because we use log customers, $\hat{\beta}_3$ is interpreted as: a 1% increase in customers is associated with a $\hat{\beta}_3 / 100$ percentage-point change in YoY PIX growth.

# Alternative Specification: Month-over-Month Growth

For robustness, we also estimate models using month-over-month (MoM) growth rates. MoM growth is noisier but uses all available observations without requiring 12 months of history.

```{r model-mom}
df_model_mom <- df_feat %>%
  select(month, g_mom_pix, selic, g_mom_ibc_br) %>%
  drop_na()

cat("MoM baseline sample:", nrow(df_model_mom), "months\n")

m_mom <- lm(g_mom_pix ~ selic + g_mom_ibc_br, data = df_model_mom)
nw_mom <- NeweyWest(m_mom, lag = max(1, floor(0.75 * nrow(df_model_mom)^(1/3))),
                    prewhite = FALSE)

cat("\n--- MoM Baseline: PIX Growth ~ Selic + IBC-Br Growth ---\n")
coeftest(m_mom, vcov = nw_mom)
```

```{r model-mom-extended}
df_model_mom_ext <- df_feat %>%
  select(month, g_mom_pix, selic, g_mom_ibc_br, ln_cust) %>%
  drop_na()

cat("MoM extended sample:", nrow(df_model_mom_ext), "months\n")

m_mom_ext <- lm(g_mom_pix ~ selic + g_mom_ibc_br + ln_cust,
                data = df_model_mom_ext)
nw_mom_ext <- NeweyWest(m_mom_ext,
                        lag = max(1, floor(0.75 * nrow(df_model_mom_ext)^(1/3))),
                        prewhite = FALSE)

cat("\n--- MoM Extended: PIX Growth ~ Selic + IBC-Br Growth + ln(Customers) ---\n")
coeftest(m_mom_ext, vcov = nw_mom_ext)
```

# Summary of Results

```{r summary-table}
extract_results <- function(model, nw_vcov, label) {
  ct <- coeftest(model, vcov = nw_vcov)
  tibble(
    Model = label,
    Term = rownames(ct),
    Estimate = ct[, "Estimate"],
    `Std. Error (HAC)` = ct[, "Std. Error"],
    `t value` = ct[, "t value"],
    `p value` = ct[, "Pr(>|t|)"],
    R2 = summary(model)$r.squared,
    N = nobs(model)
  )
}

results <- bind_rows(
  extract_results(m_base, nw_base, "Baseline (YoY)"),
  extract_results(m_ext, nw_ext, "Extended (YoY)"),
  extract_results(m_mom, nw_mom, "Baseline (MoM)"),
  extract_results(m_mom_ext, nw_mom_ext, "Extended (MoM)")
)

results %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  knitr::kable(caption = "Regression Results Summary")
```

# Caveats and Limitations

1. **Short sample**: PIX was launched in November 2020, giving us roughly 4 years of data. YoY growth calculations further reduce the usable sample to ~3 years. Customer data is available for only ~2 years.

2. **Structural break in PIX**: PIX transactions grew from zero to billions in a few years. The early explosive growth phase (2020–2022) reflects adoption dynamics (S-curve) rather than macro sensitivity. The later phase (2023–2024) may better capture steady-state macro relationships.

3. **Endogeneity**: All regressors are potentially endogenous. Selic responds to economic conditions, IBC-Br and PIX are jointly determined, and customer counts reflect the same digitization forces driving PIX. Instrumental variable or quasi-experimental approaches would be needed for causal identification.

4. **Omitted variables**: Important drivers not included: population growth, smartphone penetration, regulatory mandates (BCB requirements for PIX adoption), competitive dynamics among banks and fintechs, and seasonal patterns.

5. **Aggregation**: We use aggregate national data. Heterogeneity across regions, income groups, or institution types is not captured.

# Key R Packages

| Package | Purpose |
|---------|---------|
| `tidyverse` | Data manipulation and visualization |
| `readxl` | Reading Excel files |
| `lubridate` | Date handling |
| `scales` | Axis formatting |
| `sandwich` | HAC (Newey–West) variance estimators |
| `lmtest` | Coefficient testing with robust SEs |
| `broom` | Tidy model output |
| `zoo` | Time-series utilities |
| `gridExtra` | Arranging multiple plots |
