---
title: "Impact of Open Finance on Credit in Brazil"
subtitle: "Analyzing Credit Growth, Default Rates, PIX Adoption, and the Role of Open Finance"
author: "Pranav Mehta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

```{r include=FALSE}
required_pkgs <- c("tidyverse", "readxl", "lubridate", "scales", "sandwich", "lmtest")
to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(sandwich)
library(lmtest)
```

# Objective

This notebook examines the **impact of Open Finance on credit in Brazil**, with a focus on **financial inclusion**, by combining multiple data sources:

1. **SCR.Data** (Sistema de Informações de Crédito) — monthly aggregated credit data published by the Central Bank of Brazil (BCB), including active credit portfolios, default rates, and credit originations (flows).
2. **BCB SGS** (Sistema Gerenciador de Séries Temporais) — credit originations by modality, SME credit stock, and average interest rates/tenors.
3. **PIX transaction volumes** — Brazil's instant payment system, a core pillar of Open Finance infrastructure.
4. **Open Finance customer counts** — the number of customers with active data-sharing consents.

We investigate whether the expansion of Open Finance is associated with:

- **Credit origination growth** — higher flow of new lending (prioritized over stock measures for financial inclusion).
- **SME credit expansion** — credit access for micro, small, and medium enterprises (MPMe).
- **Consumer credit growth** — especially unsecured personal loans where information asymmetry is highest.
- **Lower default rates** — improved credit quality as information-sharing reduces adverse selection.
- **Changes in lending conditions** — whether interest rates decline as Open Finance increases competition.
- **Differential effects by client type** — whether individuals (PF) and firms (PJ) are affected differently.

# Background

## Open Finance in Brazil

Brazil's Open Finance initiative, regulated by the Central Bank, mandates that financial institutions share customer data (with consent) through standardized APIs. Launched in phases starting in 2021, the system aims to increase competition, reduce information asymmetries, and improve credit allocation. By enabling lenders to access a borrower's full financial history across institutions, Open Finance has the potential to:

- **Expand credit access** — lenders can better assess previously opaque borrowers.
- **Reduce default rates** — richer information reduces adverse selection and improves risk pricing.
- **Lower interest rates** — increased competition among lenders benefits borrowers.

## SCR (Sistema de Informações de Crédito)

The SCR is the Central Bank's credit information system, containing loan-level data from all regulated financial institutions. The public **SCR.Data** release provides monthly aggregated statistics on:

- **Carteira Ativa** (Active Portfolio) — total outstanding credit stock, computed as the sum of amounts due (to mature and overdue) across credit modalities.
- **Inadimplência** (Default Rate) — the share of the portfolio with any installment overdue by more than 90 days.
- **Ativo Problemático** (Problem Assets) — broader measure of distressed credit.

Data is disaggregated by client type (PF: individuals, PJ: firms), credit modality, state, economic sector, and other dimensions.

## PIX

PIX is Brazil's instant payment system, launched in November 2020. It enables real-time, 24/7 transfers at zero cost to end users. In the Open Finance context, PIX enables payment initiation via third-party providers — a core infrastructure pillar.

## Selic and IBC-Br

- **Selic**: Brazil's policy interest rate target, the primary instrument for influencing inflation and economic activity.
- **IBC-Br**: The Central Bank's monthly economic activity index, a high-frequency GDP proxy.

# Data Sources

## Existing Data — PIX, Customers, Macro Variables

From `data/data_monthly.xlsx`:

- **Sheet 1**: PIX, Customers, Selic, IBC-Br (January 2023 – January 2025)
- **Sheet 2**: PIX, Selic, IBC-Br (November 2020 – December 2024)

## SCR.Data — Credit Stock Variables

Monthly national aggregates from BCB's SCR.Data portal, stored locally as CSV files:

- `carteira.csv` — Active credit portfolio (R$)
- `ativo-problemtico.csv` — Problem asset rate (%)
- `inadimplncia.csv` — Default rate (%)

Data covers June 2012 to June 2025 (national totals, all client types).

## BCB SGS — Credit Originations, SME Credit, and Interest Rates

Monthly time series from BCB's SGS system, downloaded via `download_bcb_data.R`:

**Credit Originations (Concessões)** — flow-based measures of new lending:

- `concessoes_total.csv` — Total credit originations (Series 20631)
- `concessoes_pf_pessoal_total.csv` — PF personal credit total (Series 20672)
- `concessoes_pf_pessoal_nao_consig.csv` — PF personal non-payroll (Series 20666)
- `concessoes_pf_cartao.csv` — PF credit card originations (Series 20656)
- `concessoes_pj_total.csv` — PJ total originations (Series 20635)
- `concessoes_pj_capital_giro.csv` — PJ working capital (Series 20642)

**SME Credit Stock:**

- `credito_mpme.csv` — Credit balance for Micro, Small & Medium enterprises (Series 27701)

**Interest Rates:**

- `taxa_juros_pf_total.csv` — Average interest rate, PF total (Series 20716)
- `taxa_juros_pf_pessoal_nao_consig.csv` — Average monthly rate, PF personal non-payroll (Series 25464)
- `taxa_juros_pj_pequeno_porte.csv` — Average rate, PJ small businesses (Series 26429)

# Load Existing Data (PIX, Customers, Macro)

```{r load-pix-data}
df_long <- read_excel("data/data_monthly.xlsx", sheet = 2) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_short <- read_excel("data/data_monthly.xlsx", sheet = 1) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_merged <- df_long %>%
  left_join(df_short %>% select(month, cust), by = "month")

cat("PIX/Macro series:", nrow(df_long), "months,",
    as.character(min(df_long$month)), "to", as.character(max(df_long$month)), "\n")
cat("Months with customer data:", sum(!is.na(df_merged$cust)), "\n")
```

# Load SCR Credit Data from Local CSVs

The SCR credit data has been downloaded from BCB and is available as local CSV files covering June 2012 to June 2025.

## National Totals

```{r load-scr}
# Read the three SCR CSV files (national totals)
# Files use UTF-8 with BOM, comma-separated, with YYYY-MM date format
df_carteira <- read_csv("data/carteira.csv", show_col_types = FALSE) %>%
  select(date = 1, credit_portfolio = 2)

df_ativo_prob <- read_csv("data/ativo-problemtico.csv", show_col_types = FALSE) %>%
  select(date = 1, problem_asset_rate = 2)

df_inadimplencia <- read_csv("data/inadimplncia.csv", show_col_types = FALSE) %>%
  select(date = 1, default_rate = 2)

# Merge the three series by date
scr_total <- df_carteira %>%
  left_join(df_ativo_prob, by = "date") %>%
  left_join(df_inadimplencia, by = "date") %>%
  mutate(
    month = ymd(paste0(date, "-01")),
    month = ceiling_date(month, "month") - days(1)
  ) %>%
  select(month, credit_portfolio, problem_asset_rate, default_rate)

cat("SCR credit data:", nrow(scr_total), "months\n")
cat("Date range:", as.character(min(scr_total$month)),
    "to", as.character(max(scr_total$month)), "\n")
cat("Credit portfolio range: R$",
    format(min(scr_total$credit_portfolio), big.mark = ","),
    "to R$",
    format(max(scr_total$credit_portfolio), big.mark = ","), "\n")
cat("Default rate range:",
    round(min(scr_total$default_rate), 2), "% to",
    round(max(scr_total$default_rate), 2), "%\n")
summary(scr_total)
```

## PF/PJ Breakdown

```{r load-scr-pf-pj}
# Read PF/PJ breakdown files
# Columns: Date, Brasil-Todas, Brasil-Todas (dup), Brasil-PF-Todas, Brasil-PJ-Todas
df_cart_pf_pj <- read_csv("data/carteira_pf_pj.csv", show_col_types = FALSE) %>%
  select(date = 1, credit_pf = 4, credit_pj = 5)

df_ativo_pf_pj <- read_csv("data/ativo_problematico_pf_pj.csv", show_col_types = FALSE) %>%
  select(date = 1, problem_rate_pf = 4, problem_rate_pj = 5)

df_inad_pf_pj <- read_csv("data/inadimplencia_pf_pj.csv", show_col_types = FALSE) %>%
  select(date = 1, default_rate_pf = 4, default_rate_pj = 5)

scr_pf_pj <- df_cart_pf_pj %>%
  left_join(df_ativo_pf_pj, by = "date") %>%
  left_join(df_inad_pf_pj, by = "date") %>%
  mutate(
    month = ymd(paste0(date, "-01")),
    month = ceiling_date(month, "month") - days(1)
  ) %>%
  select(month, credit_pf, credit_pj, problem_rate_pf, problem_rate_pj,
         default_rate_pf, default_rate_pj)

cat("PF/PJ data:", nrow(scr_pf_pj), "months\n")
cat("PF credit range: R$", format(min(scr_pf_pj$credit_pf), big.mark = ","),
    "to R$", format(max(scr_pf_pj$credit_pf), big.mark = ","), "\n")
cat("PJ credit range: R$", format(min(scr_pf_pj$credit_pj), big.mark = ","),
    "to R$", format(max(scr_pf_pj$credit_pj), big.mark = ","), "\n")
```

## PF Credit Modalities

```{r load-scr-pf-modalities}
# Read PF modality breakdown (key sub-segments of individual credit)
# Columns: Date, Total, Total(dup), PF-Todas, PJ-Todas, then PF modalities:
#   Cartão de crédito, Consignação, Sem consignação, Habitacional,
#   Outros créditos, Rural, Veículos
df_cart_mod <- read_csv("data/carteira_pf_modalities.csv", show_col_types = FALSE) %>%
  select(date = 1,
         credit_pf_card = 6,
         credit_pf_payroll = 7,
         credit_pf_personal = 8,
         credit_pf_housing = 9,
         credit_pf_vehicles = 12)

scr_pf_mod <- df_cart_mod %>%
  mutate(
    month = ymd(paste0(date, "-01")),
    month = ceiling_date(month, "month") - days(1)
  ) %>%
  select(-date)

cat("PF modality data:", nrow(scr_pf_mod), "months\n")
cat("Key PF credit modalities (latest month, R$ Bn):\n")
latest <- scr_pf_mod %>% filter(month == max(month))
cat("  Credit card:    ", round(latest$credit_pf_card / 1e9, 1), "\n")
cat("  Payroll loans:  ", round(latest$credit_pf_payroll / 1e9, 1), "\n")
cat("  Personal loans: ", round(latest$credit_pf_personal / 1e9, 1), "\n")
cat("  Housing:        ", round(latest$credit_pf_housing / 1e9, 1), "\n")
cat("  Vehicles:       ", round(latest$credit_pf_vehicles / 1e9, 1), "\n")
```

## Credit Originations (BCB SGS)

```{r load-originations}
# BCB SGS CSV format: semicolon-separated, DD/MM/YYYY dates, comma as decimal separator
# Run download_bcb_data.R first to download these files from the BCB API

load_sgs <- function(filepath, varname) {
  if (!file.exists(filepath)) {
    cat("  MISSING:", filepath, "— run download_bcb_data.R to fetch\n")
    return(NULL)
  }
  df <- read_csv2(filepath, show_col_types = FALSE)
  names(df) <- c("date_str", "value")
  df %>%
    mutate(
      month = dmy(date_str),
      month = ceiling_date(month, "month") - days(1),
      !!varname := as.numeric(gsub(",", ".", value))
    ) %>%
    select(month, all_of(varname))
}

orig_total    <- load_sgs("data/concessoes_total.csv", "orig_total")
orig_pf_pers  <- load_sgs("data/concessoes_pf_pessoal_total.csv", "orig_pf_personal")
orig_pf_nc    <- load_sgs("data/concessoes_pf_pessoal_nao_consig.csv", "orig_pf_personal_nc")
orig_pf_card  <- load_sgs("data/concessoes_pf_cartao.csv", "orig_pf_card")
orig_pj_total <- load_sgs("data/concessoes_pj_total.csv", "orig_pj_total")
orig_pj_giro  <- load_sgs("data/concessoes_pj_capital_giro.csv", "orig_pj_working_cap")

# Merge all origination series
orig_list <- list(orig_total, orig_pf_pers, orig_pf_nc, orig_pf_card,
                  orig_pj_total, orig_pj_giro)
orig_list <- orig_list[!sapply(orig_list, is.null)]

if (length(orig_list) > 0) {
  scr_orig <- reduce(orig_list, left_join, by = "month")
  cat("Credit originations:", nrow(scr_orig), "months\n")
  cat("Date range:", as.character(min(scr_orig$month)),
      "to", as.character(max(scr_orig$month)), "\n")
} else {
  scr_orig <- tibble(month = as.Date(character()))
  cat("No origination data available. Run download_bcb_data.R first.\n")
}
```

## SME Credit Stock (BCB SGS)

```{r load-sme}
sme_credit <- load_sgs("data/credito_mpme.csv", "credit_mpme")

if (!is.null(sme_credit)) {
  cat("SME credit stock:", nrow(sme_credit), "months\n")
  cat("Date range:", as.character(min(sme_credit$month)),
      "to", as.character(max(sme_credit$month)), "\n")
} else {
  sme_credit <- tibble(month = as.Date(character()))
  cat("No SME data available. Run download_bcb_data.R first.\n")
}
```

## Interest Rates (BCB SGS)

```{r load-rates}
rate_pf_total <- load_sgs("data/taxa_juros_pf_total.csv", "rate_pf_total")
rate_pf_pers  <- load_sgs("data/taxa_juros_pf_pessoal_nao_consig.csv", "rate_pf_personal_nc")
rate_pj_small <- load_sgs("data/taxa_juros_pj_pequeno_porte.csv", "rate_pj_small")

rate_list <- list(rate_pf_total, rate_pf_pers, rate_pj_small)
rate_list <- rate_list[!sapply(rate_list, is.null)]

if (length(rate_list) > 0) {
  scr_rates <- reduce(rate_list, left_join, by = "month")
  cat("Interest rate data:", nrow(scr_rates), "months\n")
  cat("Date range:", as.character(min(scr_rates$month)),
      "to", as.character(max(scr_rates$month)), "\n")
} else {
  scr_rates <- tibble(month = as.Date(character()))
  cat("No interest rate data available. Run download_bcb_data.R first.\n")
}
```

# Merge All Data

```{r merge-all}
# Combine PIX/macro data with SCR credit data (total + PF/PJ + PF modalities)
# plus originations, SME credit, and interest rates
df_all <- df_merged %>%
  left_join(scr_total, by = "month") %>%
  left_join(scr_pf_pj, by = "month") %>%
  left_join(scr_pf_mod, by = "month")

# Add new data sources if available
if (nrow(scr_orig) > 0) df_all <- df_all %>% left_join(scr_orig, by = "month")
if (nrow(sme_credit) > 0) df_all <- df_all %>% left_join(sme_credit, by = "month")
if (nrow(scr_rates) > 0) df_all <- df_all %>% left_join(scr_rates, by = "month")

cat("Combined dataset:", nrow(df_all), "months\n")
cat("Months with credit data:", sum(!is.na(df_all$credit_portfolio)), "\n")
cat("Months with PF/PJ data:", sum(!is.na(df_all$credit_pf)), "\n")
cat("Months with customer + credit data:",
    sum(!is.na(df_all$cust) & !is.na(df_all$credit_portfolio)), "\n")
```

# Feature Engineering

```{r features}
safe_log <- function(x) ifelse(!is.na(x) & x > 0, log(x), NA_real_)

df_feat <- df_all %>%
  arrange(month) %>%
  mutate(
    ln_pix    = log(pix),
    ln_ibc_br = log(ibc_br),
    ln_cust   = safe_log(cust),
    ln_credit = safe_log(credit_portfolio),

    # PF/PJ logs
    ln_credit_pf = safe_log(credit_pf),
    ln_credit_pj = safe_log(credit_pj),

    # PF modality logs
    ln_pf_card     = safe_log(credit_pf_card),
    ln_pf_payroll  = safe_log(credit_pf_payroll),
    ln_pf_personal = safe_log(credit_pf_personal),
    ln_pf_housing  = safe_log(credit_pf_housing),

    # YoY growth rates (12-month log differences, %)
    g_yoy_ibc_br      = 100 * (ln_ibc_br - lag(ln_ibc_br, 12)),
    g_yoy_credit      = 100 * (ln_credit - lag(ln_credit, 12)),
    g_yoy_credit_pf   = 100 * (ln_credit_pf - lag(ln_credit_pf, 12)),
    g_yoy_credit_pj   = 100 * (ln_credit_pj - lag(ln_credit_pj, 12)),
    g_yoy_pf_card     = 100 * (ln_pf_card - lag(ln_pf_card, 12)),
    g_yoy_pf_payroll  = 100 * (ln_pf_payroll - lag(ln_pf_payroll, 12)),
    g_yoy_pf_personal = 100 * (ln_pf_personal - lag(ln_pf_personal, 12)),
    g_yoy_pf_housing  = 100 * (ln_pf_housing - lag(ln_pf_housing, 12)),

    # MoM growth rates
    g_mom_ibc_br = 100 * (ln_ibc_br - lag(ln_ibc_br, 1)),
    g_mom_credit = 100 * (ln_credit - lag(ln_credit, 1)),

    # Default rate YoY changes (pp)
    d_default_rate_yoy = default_rate - lag(default_rate, 12),
    d_default_pf_yoy   = default_rate_pf - lag(default_rate_pf, 12),
    d_default_pj_yoy   = default_rate_pj - lag(default_rate_pj, 12)
  )

cat("Observations with YoY credit growth:", sum(!is.na(df_feat$g_yoy_credit)), "\n")
cat("Observations with YoY PF credit growth:", sum(!is.na(df_feat$g_yoy_credit_pf)), "\n")
cat("Observations with customer data:", sum(!is.na(df_feat$ln_cust)), "\n")
```

## New Variables: Credit Originations, SME Credit, Interest Rates

```{r features-new}
# Credit origination YoY growth rates (flow-based measures)
orig_vars <- c("orig_total", "orig_pf_personal", "orig_pf_personal_nc",
               "orig_pf_card", "orig_pj_total", "orig_pj_working_cap")
existing_orig <- intersect(orig_vars, names(df_feat))

if (length(existing_orig) > 0) {
  for (v in existing_orig) {
    ln_name <- paste0("ln_", v)
    gyoy_name <- paste0("g_yoy_", v)
    df_feat <- df_feat %>%
      mutate(
        !!ln_name := safe_log(.data[[v]]),
        !!gyoy_name := 100 * (.data[[ln_name]] - lag(.data[[ln_name]], 12))
      )
  }
  cat("Origination growth variables created for:", paste(existing_orig, collapse = ", "), "\n")
  cat("Obs with YoY origination growth (total):",
      sum(!is.na(df_feat[["g_yoy_orig_total"]])), "\n")
} else {
  cat("No origination variables available.\n")
}

# SME credit YoY growth
if ("credit_mpme" %in% names(df_feat)) {
  df_feat <- df_feat %>%
    mutate(
      ln_credit_mpme = safe_log(credit_mpme),
      g_yoy_credit_mpme = 100 * (ln_credit_mpme - lag(ln_credit_mpme, 12))
    )
  cat("SME credit growth variable created. Obs:",
      sum(!is.na(df_feat$g_yoy_credit_mpme)), "\n")
} else {
  cat("No SME credit variable available.\n")
}

# Interest rate YoY changes (pp)
rate_vars <- c("rate_pf_total", "rate_pf_personal_nc", "rate_pj_small")
existing_rates <- intersect(rate_vars, names(df_feat))

if (length(existing_rates) > 0) {
  for (v in existing_rates) {
    d_name <- paste0("d_", v, "_yoy")
    df_feat <- df_feat %>%
      mutate(!!d_name := .data[[v]] - lag(.data[[v]], 12))
  }
  cat("Interest rate change variables created for:", paste(existing_rates, collapse = ", "), "\n")
} else {
  cat("No interest rate variables available.\n")
}
```

# Descriptive Analysis

## Credit Portfolio Evolution

```{r plot-credit-total}
df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_point(color = "purple4", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Total Active Credit Portfolio (SCR)",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## Credit Portfolio by Client Type

```{r plot-credit-by-type}
df_feat %>%
  filter(!is.na(credit_pf)) %>%
  select(month, Individuals = credit_pf, Firms = credit_pj) %>%
  pivot_longer(-month, names_to = "Client Type", values_to = "portfolio") %>%
  ggplot(aes(month, portfolio / 1e9, color = `Client Type`)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c("Individuals" = "steelblue", "Firms" = "darkorange")) +
  labs(title = "Active Credit Portfolio by Client Type",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## PF Credit by Modality

```{r plot-pf-modalities}
df_feat %>%
  filter(!is.na(credit_pf_card)) %>%
  select(month,
         `Credit Card` = credit_pf_card,
         `Payroll Loans` = credit_pf_payroll,
         `Personal Loans` = credit_pf_personal,
         Housing = credit_pf_housing,
         Vehicles = credit_pf_vehicles) %>%
  pivot_longer(-month, names_to = "Modality", values_to = "portfolio") %>%
  ggplot(aes(month, portfolio / 1e9, color = Modality)) +
  geom_line(linewidth = 0.9) +
  scale_y_continuous(labels = comma) +
  labs(title = "PF Credit Portfolio by Modality",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## Default Rates

```{r plot-default-rate}
df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 1) +
  labs(title = "Credit Default Rate (Inadimplência)",
       subtitle = "Share of portfolio with installments overdue > 90 days",
       x = NULL, y = "Default Rate (%)") +
  theme_minimal()
```

## Default Rate by Client Type

```{r plot-default-by-type}
df_feat %>%
  filter(!is.na(default_rate_pf)) %>%
  select(month, Individuals = default_rate_pf, Firms = default_rate_pj) %>%
  pivot_longer(-month, names_to = "Client Type", values_to = "rate") %>%
  ggplot(aes(month, rate, color = `Client Type`)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Individuals" = "steelblue", "Firms" = "darkorange")) +
  labs(title = "Default Rate by Client Type",
       x = NULL, y = "Default Rate (%)") +
  theme_minimal()
```

## Credit YoY Growth

```{r plot-credit-yoy}
df_feat %>%
  filter(!is.na(g_yoy_credit)) %>%
  ggplot(aes(month, g_yoy_credit)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Total Credit Year-over-Year Growth",
       x = NULL, y = "YoY Growth (%)") +
  theme_minimal()
```

## PIX Transaction Volumes

```{r plot-pix-levels}
df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Monthly Transaction Volume",
       subtitle = "Billions of transactions",
       x = NULL, y = "Transactions (billions)") +
  theme_minimal()
```

## Open Finance Customer Counts

```{r plot-cust}
df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "darkgreen", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Open Finance Customer Counts",
       subtitle = "Millions of customers with active data-sharing consents",
       x = NULL, y = "Customers (millions)") +
  theme_minimal()
```

## Combined Dashboard

```{r plot-dashboard, fig.height=14}
p1 <- df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "Credit Portfolio (R$ Bn)", x = NULL, y = NULL) +
  theme_minimal()

p2 <- df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 0.8) +
  labs(title = "Default Rate (%)", x = NULL, y = NULL) +
  theme_minimal()

p3 <- df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Transactions (Bn)", x = NULL, y = NULL) +
  theme_minimal()

p4 <- df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "OF Customers (Mn)", x = NULL, y = NULL) +
  theme_minimal()

p5 <- df_feat %>%
  ggplot(aes(month, selic)) +
  geom_line(color = "grey40", linewidth = 0.8) +
  labs(title = "Selic (%)", x = NULL, y = NULL) +
  theme_minimal()

p6 <- df_feat %>%
  ggplot(aes(month, ibc_br)) +
  geom_line(color = "darkorange", linewidth = 0.8) +
  labs(title = "IBC-Br", x = NULL, y = NULL) +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 1)
```

## Correlation Matrix

```{r correlation}
cor_vars <- df_feat %>%
  select(credit_portfolio, default_rate, pix, cust, selic, ibc_br) %>%
  drop_na()

cat("Correlation matrix (months with all variables, N =", nrow(cor_vars), "):\n\n")
round(cor(cor_vars), 3)
```

# Regression Framework

We estimate time-series OLS regressions with **Newey–West (HAC) standard errors** to address autocorrelation and heteroskedasticity. The analysis covers credit growth and default rates at the national level, by client type (PF/PJ), and for key PF credit modalities.

## Part A — Credit Growth Models

Tests whether Open Finance adoption is associated with **credit expansion**, controlling for macro conditions:

$$
g^{yoy}_{Credit,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 > 0$: Open Finance adoption is associated with faster credit growth.

## Part B — Default Rate Models

Tests whether Open Finance adoption is associated with **improved credit quality**:

$$
\Delta_{12} \, Default_t
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 < 0$: Open Finance adoption is associated with declining default rates (better credit quality).

# Part A: Credit Growth Models

## Baseline: Credit Growth ~ Selic + IBC-Br Growth

```{r credit-baseline}
df_m_cred <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br) %>%
  drop_na()

nw_lag <- function(n) max(1, floor(0.75 * n^(1/3)))

m_cred_base <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br, data = df_m_cred)
nw_cred_base <- NeweyWest(m_cred_base, lag = nw_lag(nrow(df_m_cred)), prewhite = FALSE)

cat("Credit baseline: N =", nrow(df_m_cred), ", R² =", round(summary(m_cred_base)$r.squared, 4), "\n")
coeftest(m_cred_base, vcov = nw_cred_base)
```

## Extended: Credit Growth ~ Selic + IBC-Br + ln(Customers)

```{r credit-extended}
df_m_cred_ext <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

m_cred_ext <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_cred_ext)
nw_cred_ext <- NeweyWest(m_cred_ext, lag = nw_lag(nrow(df_m_cred_ext)), prewhite = FALSE)

cat("Credit + OF: N =", nrow(df_m_cred_ext), ", R² =", round(summary(m_cred_ext)$r.squared, 4), "\n")
coeftest(m_cred_ext, vcov = nw_cred_ext)
```

## Credit Growth by Client Type

```{r credit-by-type}
# PF (Individuals)
df_m_pf <- df_feat %>%
  select(month, g_yoy_credit_pf, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_pf <- lm(g_yoy_credit_pf ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pf)
nw_pf <- NeweyWest(m_pf, lag = nw_lag(nrow(df_m_pf)), prewhite = FALSE)

cat("Credit PF + OF: N =", nrow(df_m_pf), ", R² =", round(summary(m_pf)$r.squared, 4), "\n")
coeftest(m_pf, vcov = nw_pf)

# PJ (Firms)
df_m_pj <- df_feat %>%
  select(month, g_yoy_credit_pj, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_pj <- lm(g_yoy_credit_pj ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pj)
nw_pj <- NeweyWest(m_pj, lag = nw_lag(nrow(df_m_pj)), prewhite = FALSE)

cat("\nCredit PJ + OF: N =", nrow(df_m_pj), ", R² =", round(summary(m_pj)$r.squared, 4), "\n")
coeftest(m_pj, vcov = nw_pj)
```

## PF Credit Growth by Modality

```{r credit-pf-modalities}
# Credit card
df_m_card <- df_feat %>%
  select(month, g_yoy_pf_card, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_card <- lm(g_yoy_pf_card ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_card)
nw_card <- NeweyWest(m_card, lag = nw_lag(nrow(df_m_card)), prewhite = FALSE)

cat("PF Card + OF: N =", nrow(df_m_card), ", R² =", round(summary(m_card)$r.squared, 4), "\n")
coeftest(m_card, vcov = nw_card)

# Personal loans (sem consignação — highest information friction)
df_m_pers <- df_feat %>%
  select(month, g_yoy_pf_personal, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_pers <- lm(g_yoy_pf_personal ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pers)
nw_pers <- NeweyWest(m_pers, lag = nw_lag(nrow(df_m_pers)), prewhite = FALSE)

cat("\nPF Personal + OF: N =", nrow(df_m_pers), ", R² =", round(summary(m_pers)$r.squared, 4), "\n")
coeftest(m_pers, vcov = nw_pers)

# Housing
df_m_house <- df_feat %>%
  select(month, g_yoy_pf_housing, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_house <- lm(g_yoy_pf_housing ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_house)
nw_house <- NeweyWest(m_house, lag = nw_lag(nrow(df_m_house)), prewhite = FALSE)

cat("\nPF Housing + OF: N =", nrow(df_m_house), ", R² =", round(summary(m_house)$r.squared, 4), "\n")
coeftest(m_house, vcov = nw_house)
```

# Part B: Default Rate Models

## Baseline: Default Rate Change ~ Selic + IBC-Br Growth

```{r default-baseline}
df_m_def <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br) %>%
  drop_na()

m_def_base <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br, data = df_m_def)
nw_def_base <- NeweyWest(m_def_base, lag = nw_lag(nrow(df_m_def)), prewhite = FALSE)

cat("Default baseline: N =", nrow(df_m_def), ", R² =", round(summary(m_def_base)$r.squared, 4), "\n")
coeftest(m_def_base, vcov = nw_def_base)
```

## Extended: Default Rate Change ~ Selic + IBC-Br + ln(Customers)

```{r default-extended}
df_m_def_ext <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

m_def_ext <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_def_ext)
nw_def_ext <- NeweyWest(m_def_ext, lag = nw_lag(nrow(df_m_def_ext)), prewhite = FALSE)

cat("Default + OF: N =", nrow(df_m_def_ext), ", R² =", round(summary(m_def_ext)$r.squared, 4), "\n")
coeftest(m_def_ext, vcov = nw_def_ext)
```

## Default Rate by Client Type

```{r default-by-type}
# PF
df_m_def_pf <- df_feat %>%
  select(month, d_default_pf_yoy, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_def_pf <- lm(d_default_pf_yoy ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_def_pf)
nw_def_pf <- NeweyWest(m_def_pf, lag = nw_lag(nrow(df_m_def_pf)), prewhite = FALSE)

cat("Default PF + OF: N =", nrow(df_m_def_pf), ", R² =", round(summary(m_def_pf)$r.squared, 4), "\n")
coeftest(m_def_pf, vcov = nw_def_pf)

# PJ
df_m_def_pj <- df_feat %>%
  select(month, d_default_pj_yoy, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

m_def_pj <- lm(d_default_pj_yoy ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_def_pj)
nw_def_pj <- NeweyWest(m_def_pj, lag = nw_lag(nrow(df_m_def_pj)), prewhite = FALSE)

cat("\nDefault PJ + OF: N =", nrow(df_m_def_pj), ", R² =", round(summary(m_def_pj)$r.squared, 4), "\n")
coeftest(m_def_pj, vcov = nw_def_pj)
```

# Part C: Credit Origination Models (Flow-Based)

Credit originations measure the **flow** of new lending, which is more directly relevant to financial inclusion than credit stock. A growing stock could reflect existing loans being maintained, while growing originations indicate new borrowers gaining access.

## Origination Growth ~ Selic + IBC-Br + ln(Customers)

```{r origination-models}
if ("g_yoy_orig_total" %in% names(df_feat)) {
  # Total originations
  df_m_orig <- df_feat %>%
    select(month, g_yoy_orig_total, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_orig) >= 10) {
    m_orig_base <- lm(g_yoy_orig_total ~ selic + g_yoy_ibc_br, data = df_m_orig)
    nw_orig_base <- NeweyWest(m_orig_base, lag = nw_lag(nrow(df_m_orig)), prewhite = FALSE)

    m_orig_ext <- lm(g_yoy_orig_total ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_orig)
    nw_orig_ext <- NeweyWest(m_orig_ext, lag = nw_lag(nrow(df_m_orig)), prewhite = FALSE)

    cat("Origination Total baseline: N =", nrow(df_m_orig),
        ", R² =", round(summary(m_orig_base)$r.squared, 4), "\n")
    coeftest(m_orig_base, vcov = nw_orig_base)
    cat("\nOrigination Total + OF: N =", nrow(df_m_orig),
        ", R² =", round(summary(m_orig_ext)$r.squared, 4), "\n")
    coeftest(m_orig_ext, vcov = nw_orig_ext)
  } else {
    cat("Insufficient observations for total origination model (N =", nrow(df_m_orig), ")\n")
  }
} else {
  cat("Origination data not available. Run download_bcb_data.R first.\n")
}
```

## PF Personal Loan Originations (Non-Payroll — Highest Information Friction)

```{r origination-pf-personal}
if ("g_yoy_orig_pf_personal_nc" %in% names(df_feat)) {
  df_m_orig_nc <- df_feat %>%
    select(month, g_yoy_orig_pf_personal_nc, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_orig_nc) >= 10) {
    m_orig_nc <- lm(g_yoy_orig_pf_personal_nc ~ selic + g_yoy_ibc_br + ln_cust,
                    data = df_m_orig_nc)
    nw_orig_nc <- NeweyWest(m_orig_nc, lag = nw_lag(nrow(df_m_orig_nc)), prewhite = FALSE)

    cat("PF Personal (non-payroll) Originations + OF: N =", nrow(df_m_orig_nc),
        ", R² =", round(summary(m_orig_nc)$r.squared, 4), "\n")
    coeftest(m_orig_nc, vcov = nw_orig_nc)
  }
}
```

## PJ Originations (SME-Relevant)

```{r origination-pj}
if ("g_yoy_orig_pj_total" %in% names(df_feat)) {
  df_m_orig_pj <- df_feat %>%
    select(month, g_yoy_orig_pj_total, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_orig_pj) >= 10) {
    m_orig_pj <- lm(g_yoy_orig_pj_total ~ selic + g_yoy_ibc_br + ln_cust,
                    data = df_m_orig_pj)
    nw_orig_pj <- NeweyWest(m_orig_pj, lag = nw_lag(nrow(df_m_orig_pj)), prewhite = FALSE)

    cat("PJ Originations + OF: N =", nrow(df_m_orig_pj),
        ", R² =", round(summary(m_orig_pj)$r.squared, 4), "\n")
    coeftest(m_orig_pj, vcov = nw_orig_pj)
  }
}

if ("g_yoy_orig_pj_working_cap" %in% names(df_feat)) {
  df_m_orig_giro <- df_feat %>%
    select(month, g_yoy_orig_pj_working_cap, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_orig_giro) >= 10) {
    m_orig_giro <- lm(g_yoy_orig_pj_working_cap ~ selic + g_yoy_ibc_br + ln_cust,
                      data = df_m_orig_giro)
    nw_orig_giro <- NeweyWest(m_orig_giro, lag = nw_lag(nrow(df_m_orig_giro)), prewhite = FALSE)

    cat("\nPJ Working Capital Originations + OF: N =", nrow(df_m_orig_giro),
        ", R² =", round(summary(m_orig_giro)$r.squared, 4), "\n")
    coeftest(m_orig_giro, vcov = nw_orig_giro)
  }
}
```

# Part D: SME Credit Models

Tests whether Open Finance adoption is associated with **credit expansion for small and medium enterprises**, a key dimension of financial inclusion.

$$
g^{yoy}_{Credit_{MPMe},t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

```{r sme-models}
if ("g_yoy_credit_mpme" %in% names(df_feat)) {
  df_m_sme <- df_feat %>%
    select(month, g_yoy_credit_mpme, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_sme) >= 10) {
    m_sme_base <- lm(g_yoy_credit_mpme ~ selic + g_yoy_ibc_br, data = df_m_sme)
    nw_sme_base <- NeweyWest(m_sme_base, lag = nw_lag(nrow(df_m_sme)), prewhite = FALSE)

    m_sme_ext <- lm(g_yoy_credit_mpme ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_sme)
    nw_sme_ext <- NeweyWest(m_sme_ext, lag = nw_lag(nrow(df_m_sme)), prewhite = FALSE)

    cat("SME Credit baseline: N =", nrow(df_m_sme),
        ", R² =", round(summary(m_sme_base)$r.squared, 4), "\n")
    coeftest(m_sme_base, vcov = nw_sme_base)
    cat("\nSME Credit + OF: N =", nrow(df_m_sme),
        ", R² =", round(summary(m_sme_ext)$r.squared, 4), "\n")
    coeftest(m_sme_ext, vcov = nw_sme_ext)
  } else {
    cat("Insufficient observations for SME model (N =", nrow(df_m_sme), ")\n")
  }
} else {
  cat("SME credit data not available. Run download_bcb_data.R first.\n")
}
```

# Part E: Interest Rate Models (Lending Conditions)

Tests whether Open Finance adoption is associated with **lower borrowing costs**, consistent with the hypothesis that data portability increases competition among lenders:

$$
\Delta_{12} \, Rate_t
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 < 0$: Open Finance is associated with declining interest rates (better lending conditions).

```{r rate-models}
# PF personal non-payroll interest rate (highest information friction segment)
if ("d_rate_pf_personal_nc_yoy" %in% names(df_feat)) {
  df_m_rate_nc <- df_feat %>%
    select(month, d_rate_pf_personal_nc_yoy, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_rate_nc) >= 10) {
    m_rate_nc_base <- lm(d_rate_pf_personal_nc_yoy ~ selic + g_yoy_ibc_br, data = df_m_rate_nc)
    nw_rate_nc_base <- NeweyWest(m_rate_nc_base, lag = nw_lag(nrow(df_m_rate_nc)), prewhite = FALSE)

    m_rate_nc <- lm(d_rate_pf_personal_nc_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                    data = df_m_rate_nc)
    nw_rate_nc <- NeweyWest(m_rate_nc, lag = nw_lag(nrow(df_m_rate_nc)), prewhite = FALSE)

    cat("PF Personal Rate baseline: N =", nrow(df_m_rate_nc),
        ", R² =", round(summary(m_rate_nc_base)$r.squared, 4), "\n")
    coeftest(m_rate_nc_base, vcov = nw_rate_nc_base)
    cat("\nPF Personal Rate + OF: N =", nrow(df_m_rate_nc),
        ", R² =", round(summary(m_rate_nc)$r.squared, 4), "\n")
    coeftest(m_rate_nc, vcov = nw_rate_nc)
  }
}

# PJ small business interest rate
if ("d_rate_pj_small_yoy" %in% names(df_feat)) {
  df_m_rate_pj <- df_feat %>%
    select(month, d_rate_pj_small_yoy, selic, g_yoy_ibc_br, ln_cust) %>% drop_na()

  if (nrow(df_m_rate_pj) >= 10) {
    m_rate_pj <- lm(d_rate_pj_small_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                    data = df_m_rate_pj)
    nw_rate_pj <- NeweyWest(m_rate_pj, lag = nw_lag(nrow(df_m_rate_pj)), prewhite = FALSE)

    cat("PJ Small Business Rate + OF: N =", nrow(df_m_rate_pj),
        ", R² =", round(summary(m_rate_pj)$r.squared, 4), "\n")
    coeftest(m_rate_pj, vcov = nw_rate_pj)
  }
}
```

# Alternative Specification: Month-over-Month Growth

For robustness, we also estimate credit models using MoM growth rates.

```{r credit-mom}
df_mom_cred <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br) %>% drop_na()

m_mom_cred <- lm(g_mom_credit ~ selic + g_mom_ibc_br, data = df_mom_cred)
nw_mom_cred <- NeweyWest(m_mom_cred, lag = nw_lag(nrow(df_mom_cred)), prewhite = FALSE)

cat("Credit MoM baseline: N =", nrow(df_mom_cred), ", R² =", round(summary(m_mom_cred)$r.squared, 4), "\n")
coeftest(m_mom_cred, vcov = nw_mom_cred)

df_mom_cred_ext <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br, ln_cust) %>% drop_na()

m_mom_cred_ext <- lm(g_mom_credit ~ selic + g_mom_ibc_br + ln_cust, data = df_mom_cred_ext)
nw_mom_cred_ext <- NeweyWest(m_mom_cred_ext, lag = nw_lag(nrow(df_mom_cred_ext)), prewhite = FALSE)

cat("\nCredit MoM + OF: N =", nrow(df_mom_cred_ext), ", R² =", round(summary(m_mom_cred_ext)$r.squared, 4), "\n")
coeftest(m_mom_cred_ext, vcov = nw_mom_cred_ext)
```

# Summary of Results

```{r summary-table}
extract_results <- function(model, nw_vcov, label) {
  ct <- coeftest(model, vcov = nw_vcov)
  tibble(
    Model = label,
    Term = rownames(ct),
    Estimate = ct[, "Estimate"],
    `SE (HAC)` = ct[, "Std. Error"],
    `t` = ct[, "t value"],
    `p` = ct[, "Pr(>|t|)"],
    R2 = summary(model)$r.squared,
    N = nobs(model)
  )
}

models <- list(
  "Credit Base (YoY)"  = list(m_cred_base, nw_cred_base),
  "Credit + OF (YoY)"  = list(m_cred_ext, nw_cred_ext),
  "Credit PF + OF"     = list(m_pf, nw_pf),
  "Credit PJ + OF"     = list(m_pj, nw_pj),
  "PF Card + OF"       = list(m_card, nw_card),
  "PF Personal + OF"   = list(m_pers, nw_pers),
  "PF Housing + OF"    = list(m_house, nw_house),
  "Default Base (YoY)" = list(m_def_base, nw_def_base),
  "Default + OF (YoY)" = list(m_def_ext, nw_def_ext),
  "Default PF + OF"    = list(m_def_pf, nw_def_pf),
  "Default PJ + OF"    = list(m_def_pj, nw_def_pj),
  "Credit Base (MoM)"  = list(m_mom_cred, nw_mom_cred),
  "Credit + OF (MoM)"  = list(m_mom_cred_ext, nw_mom_cred_ext)
)

# Add new models if they were estimated
if (exists("m_orig_base"))  models[["Orig Total Base"]]      <- list(m_orig_base, nw_orig_base)
if (exists("m_orig_ext"))   models[["Orig Total + OF"]]      <- list(m_orig_ext, nw_orig_ext)
if (exists("m_orig_nc"))    models[["Orig PF Personal + OF"]] <- list(m_orig_nc, nw_orig_nc)
if (exists("m_orig_pj"))    models[["Orig PJ + OF"]]         <- list(m_orig_pj, nw_orig_pj)
if (exists("m_orig_giro"))  models[["Orig PJ WkCap + OF"]]   <- list(m_orig_giro, nw_orig_giro)
if (exists("m_sme_base"))   models[["SME Credit Base"]]      <- list(m_sme_base, nw_sme_base)
if (exists("m_sme_ext"))    models[["SME Credit + OF"]]      <- list(m_sme_ext, nw_sme_ext)
if (exists("m_rate_nc_base")) models[["PF Rate Base"]]       <- list(m_rate_nc_base, nw_rate_nc_base)
if (exists("m_rate_nc"))    models[["PF Rate + OF"]]         <- list(m_rate_nc, nw_rate_nc)
if (exists("m_rate_pj"))    models[["PJ SmBiz Rate + OF"]]   <- list(m_rate_pj, nw_rate_pj)

map2_dfr(names(models), models, ~ extract_results(.y[[1]], .y[[2]], .x)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  knitr::kable(caption = "Regression Results Summary — Impact of Open Finance on Credit")
```

# Interpretation

## Credit Growth and Open Finance

The extended credit growth model tests whether Open Finance adoption ($\ln(\text{Customers})$) is associated with faster credit growth after controlling for the Selic rate and economic activity (IBC-Br).

**Key results:**

- **PF Personal Loans** show the strongest Open Finance association: $\hat{\beta}_{\ln(Cust)} = 7.64$ ($p = 0.003$, $R^2 = 0.56$). A 10% increase in Open Finance customers is associated with a 0.73 pp increase in YoY personal loan growth. This is the segment where information asymmetry is highest and data portability should matter most.
- **PF Housing** shows a *negative* coefficient: $\hat{\beta}_{\ln(Cust)} = -0.66$ ($p = 0.001$, $R^2 = 0.77$). Housing credit growth is slowing as Open Finance expands, but this likely reflects tightening monetary policy and the high Selic sensitivity of this modality ($\hat{\beta}_{Selic} = 0.37$, $p < 0.001$) rather than a negative Open Finance effect. Housing credit is long-duration and responds to rate cycles with a lag.

## Default Rates and Open Finance

All default rate models show a **negative and significant** coefficient on $\ln(\text{Customers})$, consistent with Open Finance improving credit quality:

| Model | $\hat{\beta}_{\ln(Cust)}$ | p-value | $R^2$ |
|-------|---------------------------|---------|-------|
| Default Total + OF | −0.449 | <0.001 | 0.959 |
| Default PF + OF | −0.185 | 0.003 | 0.935 |
| Default PJ + OF | −0.765 | 0.004 | 0.708 |

The total default rate model has the highest explanatory power ($R^2 = 0.96$), and adding $\ln(\text{Customers})$ raises $R^2$ from 0.62 (baseline) to 0.96 — a dramatic improvement. This suggests Open Finance adoption captures important variation in default rates not explained by Selic and IBC-Br alone.

## PF vs. PJ Differential Effects

Contrary to the ex-ante hypothesis that individuals would benefit more from data portability, the **PJ (firm) default rate** shows a larger Open Finance coefficient ($-0.76$) than PF ($-0.19$). This could reflect:

1. **Small and micro enterprises (MEIs)** with thin credit files benefiting from Open Finance data sharing — these are effectively "individuals" running businesses.
2. **Compositional effects**: the PJ portfolio may be more sensitive to information improvements because corporate credit decisions rely more on cash flow data that Open Finance makes portable.
3. **Confounding**: PJ defaults may be driven by pandemic recovery dynamics that happen to correlate with the Open Finance rollout period.

For credit *growth*, PF personal loans ($\hat{\beta}_3 = 7.64$) show a much stronger association than aggregate credit, consistent with Open Finance enabling lending in the highest-friction consumer segment.

## PF Modality Analysis

The modality-level results align with the information asymmetry channel:

- **Personal loans (sem consignação)**: $\hat{\beta}_3 = 7.64$, $p = 0.003$ — **strongest positive effect**. These are unsecured loans where borrower screening depends most on financial history data. Open Finance makes this data portable, enabling competition and access.
- **Housing**: $\hat{\beta}_3 = -0.66$, $p = 0.001$ — negative but likely driven by the interest rate cycle. Housing credit is the most Selic-sensitive modality, and the 2022–2024 high-Selic environment suppressed mortgage origination regardless of Open Finance.
- **Payroll loans (consignação)**: Expected to show minimal Open Finance effect since repayment is guaranteed through salary deduction — information asymmetry is already low in this segment.

## Credit Originations (Flow-Based Measures)

Credit originations (concessões) measure the flow of **new lending**, as distinct from the outstanding credit stock. This is the client's priority metric for assessing financial inclusion, as it captures whether new borrowers are gaining access to credit rather than simply whether existing loans are being maintained.

If origination data is available, the key question is whether $\hat{\beta}_3 > 0$ in the origination growth model — i.e., whether Open Finance adoption is associated with faster growth in **new** lending. The personal loan (non-payroll) segment is expected to show the strongest effect, mirroring the stock-based results.

## SME Credit and Open Finance

The SME (MPMe) credit model tests whether micro, small, and medium enterprises benefit from Open Finance. These firms often have thin credit files, making them prime beneficiaries of data portability. If $\hat{\beta}_3 > 0$, Open Finance adoption is associated with faster credit growth for SMEs. Note that the PJ default rate coefficient ($-0.76$, stronger than PF's $-0.19$) already suggests firms benefit disproportionately from Open Finance.

## Lending Conditions (Interest Rates)

The interest rate models test whether Open Finance **reduces the cost of borrowing**. The theoretical mechanism is that data portability increases competition among lenders, compressing interest rate spreads. If $\hat{\beta}_3 < 0$, Open Finance is associated with declining interest rates controlling for monetary policy (Selic) and the business cycle (IBC-Br).

## Data Gaps — Items Requiring Non-Published Data

Several client-requested analyses cannot be completed with publicly available data:

1. **Borrower segmentation** (low income, rural): Not available in public SCR aggregates. Requires BCB/Open Finance Brazil cooperation.
2. **Fintech vs. traditional FI breakdown**: SCR aggregates do not segregate by institution type. Fintech-specific lending data would enable a decomposition of consumer credit growth by institution category.
3. **OF API calls by type**: Available from the [Open Finance Dashboard](https://dashboard.openfinancebrasil.org.br/open-data/api-requests/evolution) but requires manual download. The client noted that API call volume by type "signals the direction where OF is heading to" and should be used as a treatment variable.

We recommend scheduling the introductory interview with Open Finance Brazil and the Central Bank (as the client offered) to explore the availability of these non-published data sources.

# Caveats and Limitations

1. **Short overlap period**: Open Finance customer data is available from January 2023. SCR data starts from the mid-2010s. The overlap period for models including both credit and customer data is limited.

2. **Endogeneity**: All regressors are potentially endogenous. Open Finance adoption, credit growth, and economic conditions are jointly determined. Instrumental variable or difference-in-differences approaches would be needed for causal claims.

3. **Collinearity with trend**: Customer counts are monotonically increasing and highly correlated with any time trend. Significant coefficients may capture secular trends in credit markets rather than a causal effect of Open Finance.

4. **Omitted variables**: Key omitted drivers include: regulatory changes (credit card caps, payroll lending rules), fiscal policy, housing market dynamics, commodity prices, and supply-side changes (fintech entry, bank consolidation).

5. **Aggregation**: National-level data masks heterogeneity across regions, income groups, financial institutions, and credit modalities. The impact of Open Finance may be concentrated in specific segments (e.g., low-income borrowers, consumer credit) not visible at the aggregate level.

6. **SCR data granularity**: The aggregated SCR data may group certain low-count observations under "Outros Créditos" (Other Credits), which could introduce measurement error at the modality level. National totals are less affected.

7. **Structural breaks**: The COVID-19 pandemic (2020–2021) created a major structural break in credit markets, with emergency lending programs, payment moratoria, and economic disruptions. PIX was launched during this same period, making it difficult to separate pandemic recovery effects from Open Finance effects.

# Key R Packages

| Package | Purpose |
|---------|---------|
| `tidyverse` | Data manipulation and visualization |
| `readxl` | Reading Excel files |
| `lubridate` | Date handling |
| `scales` | Axis formatting |
| `sandwich` | HAC (Newey–West) variance estimators |
| `lmtest` | Coefficient testing with robust SEs |
| `gridExtra` | Arranging multiple plots |

# Data Sources

| Source | Description | Location |
|--------|-------------|----------|
| SCR.Data — Carteira Ativa (Total) | Active credit portfolio (R$) | `data/carteira.csv` |
| SCR.Data — Carteira Ativa (PF/PJ) | Portfolio by client type | `data/carteira_pf_pj.csv` |
| SCR.Data — Carteira Ativa (PF Modalities) | PF portfolio by credit modality | `data/carteira_pf_modalities.csv` |
| SCR.Data — Ativo Problemático (Total) | Problem asset rate (%) | `data/ativo-problemtico.csv` |
| SCR.Data — Ativo Problemático (PF/PJ) | Problem rate by client type | `data/ativo_problematico_pf_pj.csv` |
| SCR.Data — Ativo Problemático (PF Modalities) | PF problem rate by modality | `data/ativo_problematico_pf_modalities.csv` |
| SCR.Data — Inadimplência (Total) | Default rate (%) | `data/inadimplncia.csv` |
| SCR.Data — Inadimplência (PF/PJ) | Default rate by client type | `data/inadimplencia_pf_pj.csv` |
| PIX / Macro / Customers | PIX volumes, Selic, IBC-Br, OF customers | `data/data_monthly.xlsx` |
| BCB SGS 20631 — Concessões Total | Total credit originations (R$) | `data/concessoes_total.csv` |
| BCB SGS 20672 — Concessões PF Pessoal | PF personal credit originations | `data/concessoes_pf_pessoal_total.csv` |
| BCB SGS 20666 — Concessões PF Não Consig | PF personal non-payroll originations | `data/concessoes_pf_pessoal_nao_consig.csv` |
| BCB SGS 20656 — Concessões PF Cartão | PF credit card originations | `data/concessoes_pf_cartao.csv` |
| BCB SGS 20635 — Concessões PJ Total | PJ total originations | `data/concessoes_pj_total.csv` |
| BCB SGS 20642 — Concessões PJ Capital Giro | PJ working capital originations | `data/concessoes_pj_capital_giro.csv` |
| BCB SGS 27701 — Crédito MPMe | SME credit stock (Micro, Small & Medium) | `data/credito_mpme.csv` |
| BCB SGS 20716 — Taxa Juros PF Total | Avg interest rate - PF total | `data/taxa_juros_pf_total.csv` |
| BCB SGS 25464 — Taxa Juros PF Pessoal NC | Avg monthly rate - PF personal non-payroll | `data/taxa_juros_pf_pessoal_nao_consig.csv` |
| BCB SGS 26429 — Taxa Juros PJ Peq Porte | Avg rate - PJ small businesses | `data/taxa_juros_pj_pequeno_porte.csv` |
