---
title: "Impact of Open Finance on Credit in Brazil"
subtitle: "Analyzing Credit Growth, Default Rates, PIX Adoption, and the Role of Open Finance"
author: "Pranav Mehta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

```{r include=FALSE}
required_pkgs <- c(
  "tidyverse", "readxl", "lubridate",
  "scales", "broom", "sandwich", "lmtest", "zoo",
  "janitor"
)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(broom)
library(sandwich)
library(lmtest)
library(zoo)
library(janitor)
```

# Objective

This notebook examines the **impact of Open Finance on credit in Brazil** by combining three data sources:

1. **SCR.Data** (Sistema de Informações de Crédito) — monthly aggregated credit data published by the Central Bank of Brazil (BCB), including active credit portfolios and default rates.
2. **PIX transaction volumes** — Brazil's instant payment system, a core pillar of Open Finance infrastructure.
3. **Open Finance customer counts** — the number of customers with active data-sharing consents.

We investigate whether the expansion of Open Finance is associated with:

- **Credit growth** — larger active credit portfolios, controlling for macro conditions.
- **Lower default rates** — improved credit quality as information-sharing reduces adverse selection.
- **Differential effects by client type** — whether individuals (PF) and firms (PJ) are affected differently.

# Background

## Open Finance in Brazil

Brazil's Open Finance initiative, regulated by the Central Bank, mandates that financial institutions share customer data (with consent) through standardized APIs. Launched in phases starting in 2021, the system aims to increase competition, reduce information asymmetries, and improve credit allocation. By enabling lenders to access a borrower's full financial history across institutions, Open Finance has the potential to:

- **Expand credit access** — lenders can better assess previously opaque borrowers.
- **Reduce default rates** — richer information reduces adverse selection and improves risk pricing.
- **Lower interest rates** — increased competition among lenders benefits borrowers.

## SCR (Sistema de Informações de Crédito)

The SCR is the Central Bank's credit information system, containing loan-level data from all regulated financial institutions. The public **SCR.Data** release provides monthly aggregated statistics on:

- **Carteira Ativa** (Active Portfolio) — total outstanding credit stock, computed as the sum of amounts due (to mature and overdue) across credit modalities.
- **Inadimplência** (Default Rate) — the share of the portfolio with any installment overdue by more than 90 days.
- **Ativo Problemático** (Problem Assets) — broader measure of distressed credit.

Data is disaggregated by client type (PF: individuals, PJ: firms), credit modality, state, economic sector, and other dimensions.

## PIX

PIX is Brazil's instant payment system, launched in November 2020. It enables real-time, 24/7 transfers at zero cost to end users. In the Open Finance context, PIX enables payment initiation via third-party providers — a core infrastructure pillar.

## Selic and IBC-Br

- **Selic**: Brazil's policy interest rate target, the primary instrument for influencing inflation and economic activity.
- **IBC-Br**: The Central Bank's monthly economic activity index, a high-frequency GDP proxy.

# Data Sources

## Existing Data — PIX, Customers, Macro Variables

From `data/data_monthly.xlsx`:

- **Sheet 1**: PIX, Customers, Selic, IBC-Br (January 2023 – January 2025)
- **Sheet 2**: PIX, Selic, IBC-Br (November 2020 – December 2024)

## SCR.Data — Credit Variables

Monthly ZIP files from BCB, available at:

```
https://www.bcb.gov.br/pda/desig/scrdata_{YEAR}.zip
```

We download data for 2020–2025 and aggregate to national monthly totals.

# Load Existing Data (PIX, Customers, Macro)

```{r load-pix-data}
df_long <- read_excel("data/data_monthly.xlsx", sheet = 2) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_short <- read_excel("data/data_monthly.xlsx", sheet = 1) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_merged <- df_long %>%
  left_join(df_short %>% select(month, cust), by = "month")

cat("PIX/Macro series:", nrow(df_long), "months,",
    as.character(min(df_long$month)), "to", as.character(max(df_long$month)), "\n")
cat("Months with customer data:", sum(!is.na(df_merged$cust)), "\n")
```

# Download and Process SCR Credit Data

```{r download-scr, cache=TRUE}
scr_dir <- "data/scr_raw"
dir.create(scr_dir, recursive = TRUE, showWarnings = FALSE)

# Download SCR ZIP files for each year
years <- 2020:2025
for (yr in years) {
  zip_path <- file.path(scr_dir, sprintf("scrdata_%d.zip", yr))
  if (!file.exists(zip_path)) {
    url <- sprintf("https://www.bcb.gov.br/pda/desig/scrdata_%d.zip", yr)
    tryCatch(
      download.file(url, zip_path, mode = "wb", quiet = TRUE),
      error = function(e) {
        cat("Could not download", yr, ":", conditionMessage(e), "\n")
      }
    )
  }
}

# List successfully downloaded files
zip_files <- list.files(scr_dir, pattern = "\\.zip$", full.names = TRUE)
cat("Downloaded ZIP files:", length(zip_files), "\n")
for (f in zip_files) cat(" ", basename(f), "-", file.size(f), "bytes\n")
```

```{r extract-scr}
# Extract all ZIP files
csv_dir <- file.path(scr_dir, "extracted")
dir.create(csv_dir, recursive = TRUE, showWarnings = FALSE)

for (zf in zip_files) {
  tryCatch(
    unzip(zf, exdir = csv_dir, overwrite = FALSE),
    error = function(e) cat("Error extracting", basename(zf), ":", conditionMessage(e), "\n")
  )
}

csv_files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)
cat("Extracted CSV files:", length(csv_files), "\n")
if (length(csv_files) > 0) cat("First file:", basename(csv_files[1]), "\n")
```

```{r inspect-scr}
# Inspect the structure of the first CSV file
# Brazilian CSVs typically use semicolons as separators and commas as decimal marks
if (length(csv_files) > 0) {
  # Read first few lines to detect format
  first_lines <- readLines(csv_files[1], n = 3, encoding = "latin1")
  cat("Header line:\n", first_lines[1], "\n\n")

  # Try reading with semicolon separator (Brazilian standard)
  scr_sample <- tryCatch(
    read_csv2(csv_files[1], n_max = 5, locale = locale(encoding = "latin1")) %>%
      clean_names(),
    error = function(e) {
      # Fallback to comma separator
      read_csv(csv_files[1], n_max = 5, locale = locale(encoding = "latin1")) %>%
        clean_names()
    }
  )

  cat("Column names:\n")
  cat(paste(names(scr_sample), collapse = ", "), "\n\n")
  glimpse(scr_sample)
}
```

```{r read-scr}
# Read all SCR CSV files and combine
# We use a function that tries semicolon first, then comma separator
read_scr_csv <- function(filepath) {
  tryCatch({
    df <- read_csv2(filepath, locale = locale(encoding = "latin1"),
                    col_types = cols(.default = "c")) %>%
      clean_names()
    df
  }, error = function(e) {
    tryCatch({
      read_csv(filepath, locale = locale(encoding = "latin1"),
               col_types = cols(.default = "c")) %>%
        clean_names()
    }, error = function(e2) {
      cat("Failed to read:", basename(filepath), "\n")
      tibble()
    })
  })
}

scr_raw <- map_dfr(csv_files, read_scr_csv, .progress = TRUE)
cat("Total SCR rows:", nrow(scr_raw), "\n")
cat("Columns:", paste(names(scr_raw), collapse = ", "), "\n")
```

```{r process-scr}
# Identify the key columns dynamically
# Expected columns (after clean_names): data_base, tipo_cliente, carteira_ativa,
# inadimplencia, ativo_problematico
scr_cols <- names(scr_raw)
cat("All SCR columns:\n")
print(scr_cols)

# Find the date column (contains "data" in name)
date_col <- scr_cols[str_detect(scr_cols, "data")][1]
# Find client type column
client_col <- scr_cols[str_detect(scr_cols, "tipo.*client|client.*tipo")][1]
# Find carteira ativa column
portfolio_col <- scr_cols[str_detect(scr_cols, "carteira|portfolio")][1]
# Find inadimplencia column
default_col <- scr_cols[str_detect(scr_cols, "inadimp|default")][1]
# Find ativo problematico column
problem_col <- scr_cols[str_detect(scr_cols, "ativo.*problem|problem.*asset")][1]

cat("\nDetected column mapping:\n")
cat("  Date:", date_col, "\n")
cat("  Client type:", client_col, "\n")
cat("  Active portfolio:", portfolio_col, "\n")
cat("  Default rate:", default_col, "\n")
cat("  Problem assets:", problem_col, "\n")
```

```{r aggregate-scr}
# Convert numeric columns (handle Brazilian decimal format: comma -> dot)
parse_br_number <- function(x) {
  as.numeric(str_replace_all(x, ",", "."))
}

# Build the aggregation dataset
scr_agg <- scr_raw %>%
  rename(
    data_base_raw = !!sym(date_col)
  )

# Conditionally rename columns that were detected
if (!is.na(client_col)) scr_agg <- scr_agg %>% rename(tipo_cliente = !!sym(client_col))
if (!is.na(portfolio_col)) scr_agg <- scr_agg %>% rename(carteira_ativa_raw = !!sym(portfolio_col))
if (!is.na(default_col)) scr_agg <- scr_agg %>% rename(inadimplencia_raw = !!sym(default_col))
if (!is.na(problem_col)) scr_agg <- scr_agg %>% rename(ativo_problematico_raw = !!sym(problem_col))

# Parse date - try multiple formats
scr_agg <- scr_agg %>%
  mutate(
    month = coalesce(
      ymd(data_base_raw),
      dmy(data_base_raw),
      my(data_base_raw),
      parse_date_time(data_base_raw, orders = c("Ym", "mY", "Ymd", "dmY"))
    ),
    month = ceiling_date(month, "month") - days(1)
  )

# Parse numeric values
if ("carteira_ativa_raw" %in% names(scr_agg)) {
  scr_agg <- scr_agg %>% mutate(carteira_ativa = parse_br_number(carteira_ativa_raw))
}
if ("inadimplencia_raw" %in% names(scr_agg)) {
  scr_agg <- scr_agg %>% mutate(inadimplencia = parse_br_number(inadimplencia_raw))
}
if ("ativo_problematico_raw" %in% names(scr_agg)) {
  scr_agg <- scr_agg %>% mutate(ativo_problematico = parse_br_number(ativo_problematico_raw))
}

cat("Date range:", as.character(min(scr_agg$month, na.rm = TRUE)),
    "to", as.character(max(scr_agg$month, na.rm = TRUE)), "\n")
cat("Rows with valid dates:", sum(!is.na(scr_agg$month)), "\n")
```

```{r scr-national-totals}
# Aggregate to national monthly totals — overall and by client type

# National total (all client types)
scr_total <- scr_agg %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(
    credit_portfolio = sum(carteira_ativa, na.rm = TRUE),
    credit_problem = sum(ativo_problematico, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    # Compute weighted default rate from portfolio-level data
    default_rate = credit_problem / credit_portfolio * 100
  )

cat("National monthly totals:", nrow(scr_total), "months\n")

# By client type (PF = individuals, PJ = firms)
if ("tipo_cliente" %in% names(scr_agg)) {
  scr_by_type <- scr_agg %>%
    filter(!is.na(month), !is.na(tipo_cliente)) %>%
    mutate(
      client_type = case_when(
        str_detect(str_to_upper(tipo_cliente), "PF|PESSOA\\s*F|INDIVIDUAL|PHYSICAL") ~ "PF",
        str_detect(str_to_upper(tipo_cliente), "PJ|PESSOA\\s*J|CORPORATE|LEGAL|JURIDIC") ~ "PJ",
        TRUE ~ "Other"
      )
    ) %>%
    filter(client_type %in% c("PF", "PJ")) %>%
    group_by(month, client_type) %>%
    summarise(
      credit_portfolio = sum(carteira_ativa, na.rm = TRUE),
      credit_problem = sum(ativo_problematico, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(default_rate = credit_problem / credit_portfolio * 100)

  # Pivot to wide format
  scr_pf <- scr_by_type %>%
    filter(client_type == "PF") %>%
    select(month,
           credit_pf = credit_portfolio,
           default_rate_pf = default_rate)

  scr_pj <- scr_by_type %>%
    filter(client_type == "PJ") %>%
    select(month,
           credit_pj = credit_portfolio,
           default_rate_pj = default_rate)

  cat("PF months:", nrow(scr_pf), " | PJ months:", nrow(scr_pj), "\n")
}

summary(scr_total)
```

# Merge All Data

```{r merge-all}
# Combine PIX/macro data with SCR credit data
df_all <- df_merged %>%
  left_join(scr_total, by = "month")

# Add PF/PJ breakdowns if available
if (exists("scr_pf")) df_all <- df_all %>% left_join(scr_pf, by = "month")
if (exists("scr_pj")) df_all <- df_all %>% left_join(scr_pj, by = "month")

cat("Combined dataset:", nrow(df_all), "months\n")
cat("Months with credit data:", sum(!is.na(df_all$credit_portfolio)), "\n")
cat("Months with customer + credit data:",
    sum(!is.na(df_all$cust) & !is.na(df_all$credit_portfolio)), "\n")
```

# Feature Engineering

```{r features}
df_feat <- df_all %>%
  arrange(month) %>%
  mutate(
    # Log transformations
    ln_pix    = log(pix),
    ln_ibc_br = log(ibc_br),
    ln_cust   = ifelse(!is.na(cust), log(cust), NA_real_),

    # Credit log transformations
    ln_credit = ifelse(!is.na(credit_portfolio) & credit_portfolio > 0,
                       log(credit_portfolio), NA_real_),

    # YoY growth rates (12-month log differences)
    g_yoy_pix      = 100 * (ln_pix - lag(ln_pix, 12)),
    g_yoy_ibc_br   = 100 * (ln_ibc_br - lag(ln_ibc_br, 12)),
    g_yoy_credit   = 100 * (ln_credit - lag(ln_credit, 12)),

    # MoM growth rates
    g_mom_pix      = 100 * (ln_pix - lag(ln_pix, 1)),
    g_mom_ibc_br   = 100 * (ln_ibc_br - lag(ln_ibc_br, 1)),
    g_mom_credit   = 100 * (ln_credit - lag(ln_credit, 1)),

    # Default rate changes
    d_default_rate     = default_rate - lag(default_rate, 1),
    d_default_rate_yoy = default_rate - lag(default_rate, 12),

    # Customer growth
    g_yoy_cust = 100 * (ln_cust - lag(ln_cust, 12)),
    g_mom_cust = 100 * (ln_cust - lag(ln_cust, 1))
  )

# PF/PJ credit features if available
if ("credit_pf" %in% names(df_feat)) {
  df_feat <- df_feat %>%
    mutate(
      ln_credit_pf     = ifelse(!is.na(credit_pf) & credit_pf > 0, log(credit_pf), NA_real_),
      ln_credit_pj     = ifelse(!is.na(credit_pj) & credit_pj > 0, log(credit_pj), NA_real_),
      g_yoy_credit_pf  = 100 * (ln_credit_pf - lag(ln_credit_pf, 12)),
      g_yoy_credit_pj  = 100 * (ln_credit_pj - lag(ln_credit_pj, 12)),
      d_default_pf_yoy = default_rate_pf - lag(default_rate_pf, 12),
      d_default_pj_yoy = default_rate_pj - lag(default_rate_pj, 12)
    )
}

cat("Observations with YoY credit growth:", sum(!is.na(df_feat$g_yoy_credit)), "\n")
cat("Observations with YoY PIX growth:", sum(!is.na(df_feat$g_yoy_pix)), "\n")
cat("Observations with customer data:", sum(!is.na(df_feat$ln_cust)), "\n")
```

# Descriptive Analysis

## Credit Portfolio Evolution

```{r plot-credit-total}
df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_point(color = "purple4", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Total Active Credit Portfolio (SCR)",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## Credit Portfolio by Client Type

```{r plot-credit-by-type, eval="credit_pf" %in% names(df_feat)}
df_feat %>%
  filter(!is.na(credit_pf)) %>%
  select(month, Individuals = credit_pf, Firms = credit_pj) %>%
  pivot_longer(-month, names_to = "Client Type", values_to = "portfolio") %>%
  ggplot(aes(month, portfolio / 1e9, color = `Client Type`)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c("Individuals" = "steelblue", "Firms" = "darkorange")) +
  labs(title = "Active Credit Portfolio by Client Type",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## Default Rates

```{r plot-default-rate}
df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 1) +
  labs(title = "Credit Default Rate (Inadimplência)",
       subtitle = "Share of portfolio with installments overdue > 90 days",
       x = NULL, y = "Default Rate (%)") +
  theme_minimal()
```

```{r plot-default-by-type, eval="default_rate_pf" %in% names(df_feat)}
df_feat %>%
  filter(!is.na(default_rate_pf)) %>%
  select(month, Individuals = default_rate_pf, Firms = default_rate_pj) %>%
  pivot_longer(-month, names_to = "Client Type", values_to = "rate") %>%
  ggplot(aes(month, rate, color = `Client Type`)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Individuals" = "steelblue", "Firms" = "darkorange")) +
  labs(title = "Default Rate by Client Type",
       x = NULL, y = "Default Rate (%)") +
  theme_minimal()
```

## Credit YoY Growth

```{r plot-credit-yoy}
df_feat %>%
  filter(!is.na(g_yoy_credit)) %>%
  ggplot(aes(month, g_yoy_credit)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Total Credit Year-over-Year Growth",
       x = NULL, y = "YoY Growth (%)") +
  theme_minimal()
```

## PIX Transaction Volumes

```{r plot-pix-levels}
df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Monthly Transaction Volume",
       subtitle = "Billions of transactions",
       x = NULL, y = "Transactions (billions)") +
  theme_minimal()
```

## Open Finance Customer Counts

```{r plot-cust}
df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "darkgreen", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Open Finance Customer Counts",
       subtitle = "Millions of customers with active data-sharing consents",
       x = NULL, y = "Customers (millions)") +
  theme_minimal()
```

## Combined Dashboard

```{r plot-dashboard, fig.height=14}
p1 <- df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "Credit Portfolio (R$ Bn)", x = NULL, y = NULL) +
  theme_minimal()

p2 <- df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 0.8) +
  labs(title = "Default Rate (%)", x = NULL, y = NULL) +
  theme_minimal()

p3 <- df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Transactions (Bn)", x = NULL, y = NULL) +
  theme_minimal()

p4 <- df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "OF Customers (Mn)", x = NULL, y = NULL) +
  theme_minimal()

p5 <- df_feat %>%
  ggplot(aes(month, selic)) +
  geom_line(color = "grey40", linewidth = 0.8) +
  labs(title = "Selic (%)", x = NULL, y = NULL) +
  theme_minimal()

p6 <- df_feat %>%
  ggplot(aes(month, ibc_br)) +
  geom_line(color = "darkorange", linewidth = 0.8) +
  labs(title = "IBC-Br", x = NULL, y = NULL) +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 1)
```

## Correlation Matrix

```{r correlation}
cor_vars <- df_feat %>%
  select(credit_portfolio, default_rate, pix, cust, selic, ibc_br) %>%
  drop_na()

if (nrow(cor_vars) >= 3) {
  cor_mat <- cor(cor_vars)
  cat("Correlation matrix (months with all variables):\n")
  cat("N =", nrow(cor_vars), "\n\n")
  round(cor_mat, 3)
} else {
  cat("Insufficient overlapping observations for full correlation matrix.\n")

  # Compute pairwise correlations for available pairs
  cor_pairs <- df_feat %>%
    select(credit_portfolio, default_rate, pix, selic, ibc_br) %>%
    drop_na()

  if (nrow(cor_pairs) >= 3) {
    cat("Pairwise correlations (credit + macro, N =", nrow(cor_pairs), "):\n\n")
    round(cor(cor_pairs), 3)
  }
}
```

# Regression Framework

We estimate time-series OLS regressions with **Newey–West (HAC) standard errors** to address autocorrelation and heteroskedasticity. The analysis proceeds in three parts:

## Part A — PIX Growth Models (Existing Analysis)

Replicates the baseline analysis: does PIX transaction growth respond to macro conditions and Open Finance adoption?

## Part B — Credit Growth Models (New)

Tests whether Open Finance adoption is associated with **credit expansion**, controlling for macro conditions:

$$
g^{yoy}_{Credit,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 > 0$: Open Finance adoption is associated with faster credit growth.

## Part C — Default Rate Models (New)

Tests whether Open Finance adoption is associated with **improved credit quality**:

$$
\Delta_{12} \, Default_t
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 < 0$: Open Finance adoption is associated with declining default rates (better credit quality).

# Part A: PIX Growth Models

## Baseline: PIX Growth ~ Selic + IBC-Br Growth

```{r pix-baseline}
df_m_pix <- df_feat %>%
  select(month, g_yoy_pix, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("PIX baseline sample:", nrow(df_m_pix), "months\n")
cat("Range:", as.character(min(df_m_pix$month)),
    "to", as.character(max(df_m_pix$month)), "\n")

m_pix_base <- lm(g_yoy_pix ~ selic + g_yoy_ibc_br, data = df_m_pix)
nw_lag <- max(1, floor(0.75 * nrow(df_m_pix)^(1/3)))
nw_pix_base <- NeweyWest(m_pix_base, lag = nw_lag, prewhite = FALSE)

cat("\n--- PIX YoY Growth ~ Selic + IBC-Br Growth ---\n")
coeftest(m_pix_base, vcov = nw_pix_base)
cat("\nR-squared:", round(summary(m_pix_base)$r.squared, 4), "\n")
```

## Extended: PIX Growth ~ Selic + IBC-Br + ln(Customers)

```{r pix-extended}
df_m_pix_ext <- df_feat %>%
  select(month, g_yoy_pix, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("PIX extended sample:", nrow(df_m_pix_ext), "months\n")

if (nrow(df_m_pix_ext) >= 5) {
  m_pix_ext <- lm(g_yoy_pix ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pix_ext)
  nw_lag_ext <- max(1, floor(0.75 * nrow(df_m_pix_ext)^(1/3)))
  nw_pix_ext <- NeweyWest(m_pix_ext, lag = nw_lag_ext, prewhite = FALSE)

  cat("\n--- PIX YoY Growth ~ Selic + IBC-Br Growth + ln(Customers) ---\n")
  coeftest(m_pix_ext, vcov = nw_pix_ext)
  cat("\nR-squared:", round(summary(m_pix_ext)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for extended PIX model.\n")
}
```

# Part B: Credit Growth Models

## Baseline: Credit Growth ~ Selic + IBC-Br Growth

```{r credit-baseline}
df_m_cred <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("Credit baseline sample:", nrow(df_m_cred), "months\n")
if (nrow(df_m_cred) > 0) {
  cat("Range:", as.character(min(df_m_cred$month)),
      "to", as.character(max(df_m_cred$month)), "\n")
}

if (nrow(df_m_cred) >= 5) {
  m_cred_base <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br, data = df_m_cred)
  nw_lag_cb <- max(1, floor(0.75 * nrow(df_m_cred)^(1/3)))
  nw_cred_base <- NeweyWest(m_cred_base, lag = nw_lag_cb, prewhite = FALSE)

  cat("\n--- Credit YoY Growth ~ Selic + IBC-Br Growth ---\n")
  coeftest(m_cred_base, vcov = nw_cred_base)
  cat("\nR-squared:", round(summary(m_cred_base)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for credit baseline model.\n")
}
```

## Extended: Credit Growth ~ Selic + IBC-Br + ln(Customers)

```{r credit-extended}
df_m_cred_ext <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("Credit extended sample:", nrow(df_m_cred_ext), "months\n")

if (nrow(df_m_cred_ext) >= 5) {
  m_cred_ext <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_cred_ext)
  nw_lag_ce <- max(1, floor(0.75 * nrow(df_m_cred_ext)^(1/3)))
  nw_cred_ext <- NeweyWest(m_cred_ext, lag = nw_lag_ce, prewhite = FALSE)

  cat("\n--- Credit YoY Growth ~ Selic + IBC-Br Growth + ln(Customers) ---\n")
  coeftest(m_cred_ext, vcov = nw_cred_ext)
  cat("\nR-squared:", round(summary(m_cred_ext)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for credit extended model.\n")
}
```

## Credit Growth by Client Type

```{r credit-by-type, eval="g_yoy_credit_pf" %in% names(df_feat)}
# PF (Individuals)
df_m_pf <- df_feat %>%
  select(month, g_yoy_credit_pf, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("PF credit extended sample:", nrow(df_m_pf), "months\n")

if (nrow(df_m_pf) >= 5) {
  m_pf <- lm(g_yoy_credit_pf ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pf)
  nw_pf <- NeweyWest(m_pf, lag = max(1, floor(0.75 * nrow(df_m_pf)^(1/3))),
                     prewhite = FALSE)

  cat("\n--- PF Credit YoY Growth ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_pf, vcov = nw_pf)
  cat("\nR-squared:", round(summary(m_pf)$r.squared, 4), "\n")
}

# PJ (Firms)
df_m_pj <- df_feat %>%
  select(month, g_yoy_credit_pj, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("\nPJ credit extended sample:", nrow(df_m_pj), "months\n")

if (nrow(df_m_pj) >= 5) {
  m_pj <- lm(g_yoy_credit_pj ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_pj)
  nw_pj <- NeweyWest(m_pj, lag = max(1, floor(0.75 * nrow(df_m_pj)^(1/3))),
                     prewhite = FALSE)

  cat("\n--- PJ Credit YoY Growth ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_pj, vcov = nw_pj)
  cat("\nR-squared:", round(summary(m_pj)$r.squared, 4), "\n")
}
```

# Part C: Default Rate Models

## Baseline: Default Rate Change ~ Selic + IBC-Br Growth

```{r default-baseline}
df_m_def <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("Default rate baseline sample:", nrow(df_m_def), "months\n")

if (nrow(df_m_def) >= 5) {
  m_def_base <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br, data = df_m_def)
  nw_lag_db <- max(1, floor(0.75 * nrow(df_m_def)^(1/3)))
  nw_def_base <- NeweyWest(m_def_base, lag = nw_lag_db, prewhite = FALSE)

  cat("\n--- YoY Change in Default Rate ~ Selic + IBC-Br Growth ---\n")
  coeftest(m_def_base, vcov = nw_def_base)
  cat("\nR-squared:", round(summary(m_def_base)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for default rate baseline model.\n")
}
```

## Extended: Default Rate Change ~ Selic + IBC-Br + ln(Customers)

```{r default-extended}
df_m_def_ext <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("Default rate extended sample:", nrow(df_m_def_ext), "months\n")

if (nrow(df_m_def_ext) >= 5) {
  m_def_ext <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                  data = df_m_def_ext)
  nw_lag_de <- max(1, floor(0.75 * nrow(df_m_def_ext)^(1/3)))
  nw_def_ext <- NeweyWest(m_def_ext, lag = nw_lag_de, prewhite = FALSE)

  cat("\n--- YoY Change in Default Rate ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_def_ext, vcov = nw_def_ext)
  cat("\nR-squared:", round(summary(m_def_ext)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for default rate extended model.\n")
}
```

## Default Rate by Client Type

```{r default-by-type, eval="d_default_pf_yoy" %in% names(df_feat)}
# PF
df_m_def_pf <- df_feat %>%
  select(month, d_default_pf_yoy, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

if (nrow(df_m_def_pf) >= 5) {
  m_def_pf <- lm(d_default_pf_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                 data = df_m_def_pf)
  nw_def_pf <- NeweyWest(m_def_pf,
                         lag = max(1, floor(0.75 * nrow(df_m_def_pf)^(1/3))),
                         prewhite = FALSE)

  cat("--- YoY Change in PF Default Rate ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_def_pf, vcov = nw_def_pf)
  cat("\nR-squared:", round(summary(m_def_pf)$r.squared, 4), "\n")
}

# PJ
df_m_def_pj <- df_feat %>%
  select(month, d_default_pj_yoy, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

if (nrow(df_m_def_pj) >= 5) {
  m_def_pj <- lm(d_default_pj_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                 data = df_m_def_pj)
  nw_def_pj <- NeweyWest(m_def_pj,
                         lag = max(1, floor(0.75 * nrow(df_m_def_pj)^(1/3))),
                         prewhite = FALSE)

  cat("\n--- YoY Change in PJ Default Rate ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_def_pj, vcov = nw_def_pj)
  cat("\nR-squared:", round(summary(m_def_pj)$r.squared, 4), "\n")
}
```

# Alternative Specification: Month-over-Month Growth

For robustness, we also estimate credit models using MoM growth rates.

```{r credit-mom}
# MoM Credit Growth ~ Selic + IBC-Br
df_mom_cred <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br) %>%
  drop_na()

cat("MoM credit baseline sample:", nrow(df_mom_cred), "months\n")

if (nrow(df_mom_cred) >= 5) {
  m_mom_cred <- lm(g_mom_credit ~ selic + g_mom_ibc_br, data = df_mom_cred)
  nw_mom_cred <- NeweyWest(m_mom_cred,
                           lag = max(1, floor(0.75 * nrow(df_mom_cred)^(1/3))),
                           prewhite = FALSE)

  cat("\n--- Credit MoM Growth ~ Selic + IBC-Br MoM Growth ---\n")
  coeftest(m_mom_cred, vcov = nw_mom_cred)
  cat("\nR-squared:", round(summary(m_mom_cred)$r.squared, 4), "\n")
}

# MoM Credit Growth ~ Selic + IBC-Br + ln(Customers)
df_mom_cred_ext <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br, ln_cust) %>%
  drop_na()

cat("\nMoM credit extended sample:", nrow(df_mom_cred_ext), "months\n")

if (nrow(df_mom_cred_ext) >= 5) {
  m_mom_cred_ext <- lm(g_mom_credit ~ selic + g_mom_ibc_br + ln_cust,
                       data = df_mom_cred_ext)
  nw_mom_cred_ext <- NeweyWest(m_mom_cred_ext,
                               lag = max(1, floor(0.75 * nrow(df_mom_cred_ext)^(1/3))),
                               prewhite = FALSE)

  cat("\n--- Credit MoM Growth ~ Selic + IBC-Br MoM + ln(Customers) ---\n")
  coeftest(m_mom_cred_ext, vcov = nw_mom_cred_ext)
  cat("\nR-squared:", round(summary(m_mom_cred_ext)$r.squared, 4), "\n")
}
```

```{r pix-mom}
# MoM PIX models (from original analysis)
df_mom_pix <- df_feat %>%
  select(month, g_mom_pix, selic, g_mom_ibc_br) %>%
  drop_na()

m_mom_pix <- lm(g_mom_pix ~ selic + g_mom_ibc_br, data = df_mom_pix)
nw_mom_pix <- NeweyWest(m_mom_pix,
                        lag = max(1, floor(0.75 * nrow(df_mom_pix)^(1/3))),
                        prewhite = FALSE)

cat("--- PIX MoM Growth ~ Selic + IBC-Br MoM ---\n")
coeftest(m_mom_pix, vcov = nw_mom_pix)

# Extended
df_mom_pix_ext <- df_feat %>%
  select(month, g_mom_pix, selic, g_mom_ibc_br, ln_cust) %>%
  drop_na()

if (nrow(df_mom_pix_ext) >= 5) {
  m_mom_pix_ext <- lm(g_mom_pix ~ selic + g_mom_ibc_br + ln_cust,
                      data = df_mom_pix_ext)
  nw_mom_pix_ext <- NeweyWest(m_mom_pix_ext,
                              lag = max(1, floor(0.75 * nrow(df_mom_pix_ext)^(1/3))),
                              prewhite = FALSE)

  cat("\n--- PIX MoM Growth ~ Selic + IBC-Br MoM + ln(Customers) ---\n")
  coeftest(m_mom_pix_ext, vcov = nw_mom_pix_ext)
}
```

# Summary of Results

```{r summary-table}
extract_results <- function(model, nw_vcov, label) {
  ct <- coeftest(model, vcov = nw_vcov)
  tibble(
    Model = label,
    Term = rownames(ct),
    Estimate = ct[, "Estimate"],
    `Std. Error (HAC)` = ct[, "Std. Error"],
    `t value` = ct[, "t value"],
    `p value` = ct[, "Pr(>|t|)"],
    R2 = summary(model)$r.squared,
    N = nobs(model)
  )
}

# Collect all estimated models
results_list <- list()

# PIX models (always available)
results_list[["PIX Base (YoY)"]] <- extract_results(m_pix_base, nw_pix_base, "PIX Base (YoY)")
if (exists("m_pix_ext")) {
  results_list[["PIX + OF (YoY)"]] <- extract_results(m_pix_ext, nw_pix_ext, "PIX + OF (YoY)")
}
results_list[["PIX Base (MoM)"]] <- extract_results(m_mom_pix, nw_mom_pix, "PIX Base (MoM)")
if (exists("m_mom_pix_ext")) {
  results_list[["PIX + OF (MoM)"]] <- extract_results(m_mom_pix_ext, nw_mom_pix_ext, "PIX + OF (MoM)")
}

# Credit models
if (exists("m_cred_base")) {
  results_list[["Credit Base (YoY)"]] <- extract_results(m_cred_base, nw_cred_base, "Credit Base (YoY)")
}
if (exists("m_cred_ext")) {
  results_list[["Credit + OF (YoY)"]] <- extract_results(m_cred_ext, nw_cred_ext, "Credit + OF (YoY)")
}
if (exists("m_mom_cred")) {
  results_list[["Credit Base (MoM)"]] <- extract_results(m_mom_cred, nw_mom_cred, "Credit Base (MoM)")
}
if (exists("m_mom_cred_ext")) {
  results_list[["Credit + OF (MoM)"]] <- extract_results(m_mom_cred_ext, nw_mom_cred_ext, "Credit + OF (MoM)")
}

# Default rate models
if (exists("m_def_base")) {
  results_list[["Default Base (YoY)"]] <- extract_results(m_def_base, nw_def_base, "Default Base (YoY)")
}
if (exists("m_def_ext")) {
  results_list[["Default + OF (YoY)"]] <- extract_results(m_def_ext, nw_def_ext, "Default + OF (YoY)")
}

# PF/PJ models
if (exists("m_pf")) {
  results_list[["Credit PF + OF"]] <- extract_results(m_pf, nw_pf, "Credit PF + OF")
}
if (exists("m_pj")) {
  results_list[["Credit PJ + OF"]] <- extract_results(m_pj, nw_pj, "Credit PJ + OF")
}
if (exists("m_def_pf")) {
  results_list[["Default PF + OF"]] <- extract_results(m_def_pf, nw_def_pf, "Default PF + OF")
}
if (exists("m_def_pj")) {
  results_list[["Default PJ + OF"]] <- extract_results(m_def_pj, nw_def_pj, "Default PJ + OF")
}

results <- bind_rows(results_list)

results %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  knitr::kable(caption = "Full Regression Results Summary — Impact of Open Finance on Credit")
```

# Interpretation

## Credit Growth and Open Finance

The extended credit growth model tests whether Open Finance adoption ($\ln(\text{Customers})$) is associated with faster credit growth after controlling for the Selic rate and economic activity (IBC-Br). A positive and significant coefficient on $\ln(\text{Customers})$ would suggest that Open Finance is associated with credit expansion — consistent with the hypothesis that data-sharing reduces information asymmetries and enables broader lending.

Key considerations:

- **Selic**: Higher interest rates are expected to dampen credit growth ($\hat{\beta}_1 < 0$) as borrowing costs increase and both supply and demand contract.
- **IBC-Br**: Stronger economic activity should support credit growth ($\hat{\beta}_2 > 0$) through both higher demand for credit and improved creditworthiness.
- **Open Finance adoption**: A positive $\hat{\beta}_3$ is consistent with Open Finance enabling credit expansion, but could also reflect concurrent digitization trends.

## Default Rates and Open Finance

The default rate model tests whether Open Finance adoption is associated with declining default rates. A negative coefficient on $\ln(\text{Customers})$ would suggest that information-sharing improves credit quality — consistent with the hypothesis that better data allows lenders to screen borrowers more effectively.

Key considerations:

- Declining default rates during a period of Open Finance expansion could reflect improved risk assessment (causal) or simply economic recovery reducing distress (confounding).
- The PF (individual) vs. PJ (firm) breakdown helps distinguish channels: individuals may benefit more from data portability if their credit histories were previously fragmented across institutions.

## PF vs. PJ Differential Effects

Open Finance is expected to have a stronger impact on **individual (PF) credit** because:

1. Individuals are more likely to have fragmented financial relationships across multiple institutions.
2. Data portability enables "thin-file" borrowers to demonstrate creditworthiness using transaction history.
3. Firms already have established credit relationships and more standardized financial reporting.

A larger positive $\hat{\beta}_3$ for PF credit growth (or a more negative $\hat{\beta}_3$ for PF default rates) relative to PJ would be consistent with this channel.

# Caveats and Limitations

1. **Short overlap period**: Open Finance customer data is available from January 2023. SCR data starts from the mid-2010s. The overlap period for models including both credit and customer data is limited.

2. **Endogeneity**: All regressors are potentially endogenous. Open Finance adoption, credit growth, and economic conditions are jointly determined. Instrumental variable or difference-in-differences approaches would be needed for causal claims.

3. **Collinearity with trend**: Customer counts are monotonically increasing and highly correlated with any time trend. Significant coefficients may capture secular trends in credit markets rather than a causal effect of Open Finance.

4. **Omitted variables**: Key omitted drivers include: regulatory changes (credit card caps, payroll lending rules), fiscal policy, housing market dynamics, commodity prices, and supply-side changes (fintech entry, bank consolidation).

5. **Aggregation**: National-level data masks heterogeneity across regions, income groups, financial institutions, and credit modalities. The impact of Open Finance may be concentrated in specific segments (e.g., low-income borrowers, consumer credit) not visible at the aggregate level.

6. **SCR data granularity**: The aggregated SCR data may group certain low-count observations under "Outros Créditos" (Other Credits), which could introduce measurement error at the modality level. National totals are less affected.

7. **Structural breaks**: The COVID-19 pandemic (2020–2021) created a major structural break in credit markets, with emergency lending programs, payment moratoria, and economic disruptions. PIX was launched during this same period, making it difficult to separate pandemic recovery effects from Open Finance effects.

# Key R Packages

| Package | Purpose |
|---------|---------|
| `tidyverse` | Data manipulation and visualization |
| `readxl` | Reading Excel files |
| `lubridate` | Date handling |
| `scales` | Axis formatting |
| `sandwich` | HAC (Newey–West) variance estimators |
| `lmtest` | Coefficient testing with robust SEs |
| `broom` | Tidy model output |
| `zoo` | Time-series utilities |
| `janitor` | Column name cleaning |
| `gridExtra` | Arranging multiple plots |

# Data Sources

| Source | Description | URL |
|--------|-------------|-----|
| SCR.Data | Monthly credit aggregates from BCB | `https://www.bcb.gov.br/pda/desig/scrdata_{YEAR}.zip` |
| SCR.Data Methodology | Full methodology documentation | `https://www.bcb.gov.br/content/estabilidadefinanceira/scr/scr.data/scr_data_metodologia.pdf` |
| Open Data Portal | BCB open data | `https://dadosabertos.bcb.gov.br/dataset/scr_data` |
