---
title: "Open Finance (Brazil) — Quantitative Analysis (Credit Focus)"
subtitle: "BCB Open Data (SGS/BCData) + optional IF.data (institution-level) extension"
author: "Pranav Mehta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

```{r include=FALSE}
required_pkgs <- c(
  "tidyverse", "httr", "jsonlite", "lubridate",
  "scales", "broom", "sandwich", "lmtest", "zoo"
)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(httr)
library(jsonlite)
library(lubridate)
library(scales)
library(broom)
library(sandwich)
library(lmtest)
library(zoo)
```

# Objective (Credit Focus)

This notebook quantifies and analyzes credit-market outcomes in Brazil as a proxy for the credit segment of the Open Finance ecosystem, using publicly available Central Bank of Brazil (BCB) SGS/BCData time series.

Because Open Finance “credit outcomes” are not directly observed in public BCB data (e.g., we do not observe loan-level Open Finance originations), we use aggregate credit metrics (credit stocks and their growth rates) to characterize the broader credit cycle and provide a baseline to which Open Finance adoption proxies (e.g., consent counts or API activity) can later be added.

# What is Selic?

Selic is Brazil’s policy interest rate target. It is the primary instrument used by the Central Bank of Brazil to influence inflation and economic activity. In macro-finance terms:

- When Selic rises, borrowing costs increase and credit growth often slows (typically with lags).
- When Selic falls, credit conditions usually loosen and credit growth may accelerate.

In this notebook, Selic is treated as a policy/control variable to help separate broad macro forces (e.g., GDP growth) from monetary conditions.

# Data Sources and Series Codes

We use the following BCB SGS series (monthly or monthly-aggregated):

- credit_total_stock (20539): total credit outstanding (stock)
- credit_pj_stock (20540): corporate credit outstanding (PJ)
- credit_pf_free (20570): household credit outstanding (PF, free resources)
- selic_target (4189): Selic policy rate
- gdp_nominal_month (4380): monthly nominal GDP proxy

```{r series-map}
series_map <- c(
  credit_total_stock = 20539,
  credit_pj_stock    = 20540,
  credit_pf_free     = 20570,
  selic_target       = 4189,   # working Selic series
  gdp_nominal_month  = 4380
)

series_map
```

# Download Helper (SGS/BCData API)


BCB provides time series via the SGS/BCData API:

> https://api.bcb.gov.br/dados/serie/bcdata.sgs.{code}/dados?formato=json&dataInicial=dd/mm/yyyy&dataFinal=dd/mm/yyyy

```{r sgs-function}
get_sgs_series <- function(code,
                           start = "01/01/2013",
                           end   = format(Sys.Date(), "%d/%m/%Y")) {

  base <- paste0("https://api.bcb.gov.br/dados/serie/bcdata.sgs.", code, "/dados")

  resp <- httr::GET(
    url = base,
    query = list(
      formato     = "json",
      dataInicial = start,
      dataFinal   = end
    ),
    httr::add_headers(`User-Agent` = "R (Open Finance Capstone)")
  )

  httr::stop_for_status(resp)

  raw <- jsonlite::fromJSON(httr::content(resp, as = "text", encoding = "UTF-8"))

  if (length(raw) == 0) {
    stop(paste("Series", code, "returned 0 rows."))
  }

  tibble(
    date  = lubridate::dmy(raw$data),
    value = readr::parse_number(as.character(raw$valor))
  ) %>%
    arrange(date)
}
```

# Download and Construct Monthly Panel

```{r download}
start_date <- "01/01/2013"
end_date   <- format(Sys.Date(), "%d/%m/%Y")

df_list <- purrr::imap(
  series_map,
  ~ get_sgs_series(.x, start = start_date, end = end_date) %>%
      mutate(series = .y)
)

df_long <- bind_rows(df_list)

df_long %>% count(series)
```

## Align to month-end and pivot wide

Different SGS series sometimes have different within-month timestamps (e.g., some appear as month-start). To align consistently, we:

- Convert each observation date to a month-end index
- Take the last observation in that month for each series
- Pivot into a single monthly dataset

```{r monthly-panel}
df_monthly <- df_long %>%
  mutate(month = ceiling_date(date, "month") - days(1)) %>%
  group_by(series, month) %>%
  summarise(value = last(na.omit(value)), .groups="drop") %>%
  pivot_wider(names_from = series, values_from = value) %>%
  arrange(month)

names(df_monthly)
```

# Feature Engineering

Credit series here are stocks (levels). For interpretability in regressions, we compute:

- logs of credit and GDP
- year-over-year (YoY) growth in credit and GDP using log differences:

\[
g^{yoy}_t \approx 100 \times \left(\ln X_t - \ln X_{t-12}\right)
\]


```{r features}
df_feat <- df_monthly %>%
  mutate(
    ln_credit_total = log(credit_total_stock),
    ln_credit_pj    = log(credit_pj_stock),
    ln_credit_pf    = log(credit_pf_free),
    ln_gdp          = log(gdp_nominal_month),

    g_yoy_total = 100 * (ln_credit_total - lag(ln_credit_total, 12)),
    g_yoy_pj    = 100 * (ln_credit_pj - lag(ln_credit_pj, 12)),
    g_yoy_pf    = 100 * (ln_credit_pf - lag(ln_credit_pf, 12)),
    gdp_yoy     = 100 * (ln_gdp - lag(ln_gdp, 12))
  )
```

# Descriptive Plots

## Credit Stocks

```{r plot-levels}
df_feat %>%
  select(month, credit_total_stock, credit_pj_stock, credit_pf_free) %>%
  pivot_longer(-month, names_to="series", values_to="value") %>%
  ggplot(aes(month, value)) +
  geom_line() +
  facet_wrap(~series, scales="free_y", ncol=1) +
  scale_y_continuous(labels = comma) +
  labs(title="Credit Stocks in Brazil",
       x=NULL, y=NULL) +
  theme_minimal()
```

## YoY Growth

```{r plot-yoy}
df_feat %>%
  select(month, g_yoy_total, g_yoy_pj, g_yoy_pf) %>%
  pivot_longer(-month, names_to="series", values_to="yoy") %>%
  ggplot(aes(month, yoy)) +
  geom_line() +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~series, ncol=1) +
  labs(title="YoY Credit Growth",
       x=NULL, y="Percent") +
  theme_minimal()
```
# Regression Framework

We estimate time-series OLS regressions where the dependent variable is **year-over-year (YoY) credit growth**, and regressors include:

- Selic (policy rate)  
- YoY nominal GDP growth (macro activity)  

## Baseline Regression Equation

For segment $s \in \{\text{total}, \text{PF}, \text{PJ}\}$:

\[
g^{yoy}_{credit,s,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{GDP,t}
+
\varepsilon_t
\]

where:

- $g^{yoy}_{credit,s,t}$ is YoY credit growth for segment $s$ at time $t$  
- $Selic_t$ is the policy interest rate  
- $g^{yoy}_{GDP,t}$ is YoY nominal GDP growth  
- $\varepsilon_t$ is the error term  

## Interpretation

- $\beta_1$ captures the association between monetary conditions and credit growth.  
- $\beta_2$ captures **procyclicality**: whether credit grows faster during stronger macroeconomic conditions.  

## OLS with Newey–West SE

Monthly time-series data commonly exhibit:

- Autocorrelation (serial correlation in residuals)  
- Heteroskedasticity (non-constant variance of errors)  

To address this, we use **Newey–West (HAC)** standard errors with a 12-month lag.

This approach:

- Avoids overconfident inference due to serial correlation  
- Retains a simple, interpretable regression structure  
- Provides consistent standard errors under fairly general time-series conditions  

# Modeling Dataset

```{r model-data}
df_model <- df_feat %>%
  select(month,
         g_yoy_total, g_yoy_pf, g_yoy_pj,
         selic_target,
         gdp_yoy) %>%
  drop_na()

summary(df_model)
```

# Results: Total Credit Model (HAC SE)

## Total Credit Growth

```{r reg-total}
m_total <- lm(g_yoy_total ~ selic_target + gdp_yoy, data = df_model)

nw_total <- NeweyWest(m_total, lag = 12, prewhite = FALSE)

coeftest(m_total, vcov = nw_total)
```
## Interpretation (Total)

In the current specification:

- GDP YoY is positive and statistically significant (in your run: ~0.60, p < 0.01), suggesting credit growth is procyclical.
- Selic is negative but statistically insignificant, meaning we do not detect a stable contemporaneous linear association between Selic and YoY credit growth in this model.

This is not unexpected because:

- monetary policy effects often operate with lags
- Selic is endogenous (rises when the economy is strong), which can blur its relationship with credit
- credit stocks are slow-moving compared to credit flows

## Results by Segment: PF vs PJ (HAC SE)

```{r reg-segments}
m_pf <- lm(g_yoy_pf ~ selic_target + gdp_yoy, data = df_model)
m_pj <- lm(g_yoy_pj ~ selic_target + gdp_yoy, data = df_model)

nw_pf <- NeweyWest(m_pf, lag=12)
nw_pj <- NeweyWest(m_pj, lag=12)

list(
  PF = coeftest(m_pf, vcov=nw_pf),
  PJ = coeftest(m_pj, vcov=nw_pj)
)
```

## Interpretation

- PF (households): GDP effect tends to be positive and sometimes closer to conventional significance; Selic remains insignificant.
- PJ (corporates): estimates are less precise (larger SE), likely reflecting more volatile corporate credit dynamics and the coarse nature of aggregate series.
