---
title: "Impact of Open Finance on Credit in Brazil"
subtitle: "Analyzing Credit Growth, Default Rates, PIX Adoption, and the Role of Open Finance"
author: "Pranav Mehta"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

```{r include=FALSE}
required_pkgs <- c(
  "tidyverse", "readxl", "lubridate",
  "scales", "broom", "sandwich", "lmtest", "zoo"
)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(broom)
library(sandwich)
library(lmtest)
library(zoo)
```

# Objective

This notebook examines the **impact of Open Finance on credit in Brazil** by combining three data sources:

1. **SCR.Data** (Sistema de Informações de Crédito) — monthly aggregated credit data published by the Central Bank of Brazil (BCB), including active credit portfolios and default rates.
2. **PIX transaction volumes** — Brazil's instant payment system, a core pillar of Open Finance infrastructure.
3. **Open Finance customer counts** — the number of customers with active data-sharing consents.

We investigate whether the expansion of Open Finance is associated with:

- **Credit growth** — larger active credit portfolios, controlling for macro conditions.
- **Lower default rates** — improved credit quality as information-sharing reduces adverse selection.
- **Differential effects by client type** — whether individuals (PF) and firms (PJ) are affected differently.

# Background

## Open Finance in Brazil

Brazil's Open Finance initiative, regulated by the Central Bank, mandates that financial institutions share customer data (with consent) through standardized APIs. Launched in phases starting in 2021, the system aims to increase competition, reduce information asymmetries, and improve credit allocation. By enabling lenders to access a borrower's full financial history across institutions, Open Finance has the potential to:

- **Expand credit access** — lenders can better assess previously opaque borrowers.
- **Reduce default rates** — richer information reduces adverse selection and improves risk pricing.
- **Lower interest rates** — increased competition among lenders benefits borrowers.

## SCR (Sistema de Informações de Crédito)

The SCR is the Central Bank's credit information system, containing loan-level data from all regulated financial institutions. The public **SCR.Data** release provides monthly aggregated statistics on:

- **Carteira Ativa** (Active Portfolio) — total outstanding credit stock, computed as the sum of amounts due (to mature and overdue) across credit modalities.
- **Inadimplência** (Default Rate) — the share of the portfolio with any installment overdue by more than 90 days.
- **Ativo Problemático** (Problem Assets) — broader measure of distressed credit.

Data is disaggregated by client type (PF: individuals, PJ: firms), credit modality, state, economic sector, and other dimensions.

## PIX

PIX is Brazil's instant payment system, launched in November 2020. It enables real-time, 24/7 transfers at zero cost to end users. In the Open Finance context, PIX enables payment initiation via third-party providers — a core infrastructure pillar.

## Selic and IBC-Br

- **Selic**: Brazil's policy interest rate target, the primary instrument for influencing inflation and economic activity.
- **IBC-Br**: The Central Bank's monthly economic activity index, a high-frequency GDP proxy.

# Data Sources

## Existing Data — PIX, Customers, Macro Variables

From `data/data_monthly.xlsx`:

- **Sheet 1**: PIX, Customers, Selic, IBC-Br (January 2023 – January 2025)
- **Sheet 2**: PIX, Selic, IBC-Br (November 2020 – December 2024)

## SCR.Data — Credit Variables

Monthly national aggregates from BCB's SCR.Data portal, stored locally as CSV files:

- `carteira.csv` — Active credit portfolio (R$)
- `ativo-problemtico.csv` — Problem asset rate (%)
- `inadimplncia.csv` — Default rate (%)

Data covers June 2012 to June 2025 (national totals, all client types).

# Load Existing Data (PIX, Customers, Macro)

```{r load-pix-data}
df_long <- read_excel("data/data_monthly.xlsx", sheet = 2) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_short <- read_excel("data/data_monthly.xlsx", sheet = 1) %>%
  mutate(month = dmy(paste0("01/", month)),
         month = ceiling_date(month, "month") - days(1))

df_merged <- df_long %>%
  left_join(df_short %>% select(month, cust), by = "month")

cat("PIX/Macro series:", nrow(df_long), "months,",
    as.character(min(df_long$month)), "to", as.character(max(df_long$month)), "\n")
cat("Months with customer data:", sum(!is.na(df_merged$cust)), "\n")
```

# Load SCR Credit Data from Local CSVs

The SCR credit data has been downloaded from BCB and is available as local CSV files covering June 2012 to June 2025. These are national aggregates ("Brasil - Todas") for all client types combined.

```{r load-scr}
# Read the three SCR CSV files (national totals)
# Files use UTF-8 with BOM, comma-separated, with YYYY-MM date format
df_carteira <- read_csv("data/carteira.csv", show_col_types = FALSE) %>%
  select(date = 1, credit_portfolio = 2)

df_ativo_prob <- read_csv("data/ativo-problemtico.csv", show_col_types = FALSE) %>%
  select(date = 1, problem_asset_rate = 2)

df_inadimplencia <- read_csv("data/inadimplncia.csv", show_col_types = FALSE) %>%
  select(date = 1, default_rate = 2)

# Merge the three series by date
scr_total <- df_carteira %>%
  left_join(df_ativo_prob, by = "date") %>%
  left_join(df_inadimplencia, by = "date") %>%
  mutate(
    month = ymd(paste0(date, "-01")),
    month = ceiling_date(month, "month") - days(1)
  ) %>%
  select(month, credit_portfolio, problem_asset_rate, default_rate)

cat("SCR credit data:", nrow(scr_total), "months\n")
cat("Date range:", as.character(min(scr_total$month)),
    "to", as.character(max(scr_total$month)), "\n")
cat("Credit portfolio range: R$",
    format(min(scr_total$credit_portfolio), big.mark = ","),
    "to R$",
    format(max(scr_total$credit_portfolio), big.mark = ","), "\n")
cat("Default rate range:",
    round(min(scr_total$default_rate), 2), "% to",
    round(max(scr_total$default_rate), 2), "%\n")
summary(scr_total)
```

# Merge All Data

```{r merge-all}
# Combine PIX/macro data with SCR credit data
df_all <- df_merged %>%
  left_join(scr_total, by = "month")

cat("Combined dataset:", nrow(df_all), "months\n")
cat("Months with credit data:", sum(!is.na(df_all$credit_portfolio)), "\n")
cat("Months with customer + credit data:",
    sum(!is.na(df_all$cust) & !is.na(df_all$credit_portfolio)), "\n")
```

# Feature Engineering

```{r features}
df_feat <- df_all %>%
  arrange(month) %>%
  mutate(
    # Log transformations
    ln_pix    = log(pix),
    ln_ibc_br = log(ibc_br),
    ln_cust   = ifelse(!is.na(cust), log(cust), NA_real_),

    # Credit log transformations
    ln_credit = ifelse(!is.na(credit_portfolio) & credit_portfolio > 0,
                       log(credit_portfolio), NA_real_),

    # YoY growth rates (12-month log differences)
    g_yoy_pix      = 100 * (ln_pix - lag(ln_pix, 12)),
    g_yoy_ibc_br   = 100 * (ln_ibc_br - lag(ln_ibc_br, 12)),
    g_yoy_credit   = 100 * (ln_credit - lag(ln_credit, 12)),

    # MoM growth rates
    g_mom_pix      = 100 * (ln_pix - lag(ln_pix, 1)),
    g_mom_ibc_br   = 100 * (ln_ibc_br - lag(ln_ibc_br, 1)),
    g_mom_credit   = 100 * (ln_credit - lag(ln_credit, 1)),

    # Default rate changes
    d_default_rate     = default_rate - lag(default_rate, 1),
    d_default_rate_yoy = default_rate - lag(default_rate, 12),

    # Customer growth
    g_yoy_cust = 100 * (ln_cust - lag(ln_cust, 12)),
    g_mom_cust = 100 * (ln_cust - lag(ln_cust, 1))
  )

cat("Observations with YoY credit growth:", sum(!is.na(df_feat$g_yoy_credit)), "\n")
cat("Observations with YoY PIX growth:", sum(!is.na(df_feat$g_yoy_pix)), "\n")
cat("Observations with customer data:", sum(!is.na(df_feat$ln_cust)), "\n")
```

# Descriptive Analysis

## Credit Portfolio Evolution

```{r plot-credit-total}
df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_point(color = "purple4", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Total Active Credit Portfolio (SCR)",
       subtitle = "R$ billions",
       x = NULL, y = "Credit Portfolio (R$ Bn)") +
  theme_minimal()
```

## Default Rates

```{r plot-default-rate}
df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 1) +
  labs(title = "Credit Default Rate (Inadimplência)",
       subtitle = "Share of portfolio with installments overdue > 90 days",
       x = NULL, y = "Default Rate (%)") +
  theme_minimal()
```

## Credit YoY Growth

```{r plot-credit-yoy}
df_feat %>%
  filter(!is.na(g_yoy_credit)) %>%
  ggplot(aes(month, g_yoy_credit)) +
  geom_line(color = "purple4", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Total Credit Year-over-Year Growth",
       x = NULL, y = "YoY Growth (%)") +
  theme_minimal()
```

## PIX Transaction Volumes

```{r plot-pix-levels}
df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Monthly Transaction Volume",
       subtitle = "Billions of transactions",
       x = NULL, y = "Transactions (billions)") +
  theme_minimal()
```

## Open Finance Customer Counts

```{r plot-cust}
df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "darkgreen", size = 1.2) +
  scale_y_continuous(labels = comma) +
  labs(title = "Open Finance Customer Counts",
       subtitle = "Millions of customers with active data-sharing consents",
       x = NULL, y = "Customers (millions)") +
  theme_minimal()
```

## Combined Dashboard

```{r plot-dashboard, fig.height=14}
p1 <- df_feat %>%
  filter(!is.na(credit_portfolio)) %>%
  ggplot(aes(month, credit_portfolio / 1e9)) +
  geom_line(color = "purple4", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "Credit Portfolio (R$ Bn)", x = NULL, y = NULL) +
  theme_minimal()

p2 <- df_feat %>%
  filter(!is.na(default_rate)) %>%
  ggplot(aes(month, default_rate)) +
  geom_line(color = "firebrick", linewidth = 0.8) +
  labs(title = "Default Rate (%)", x = NULL, y = NULL) +
  theme_minimal()

p3 <- df_feat %>%
  ggplot(aes(month, pix / 1e9)) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "PIX Transactions (Bn)", x = NULL, y = NULL) +
  theme_minimal()

p4 <- df_feat %>%
  filter(!is.na(cust)) %>%
  ggplot(aes(month, cust / 1e6)) +
  geom_line(color = "darkgreen", linewidth = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "OF Customers (Mn)", x = NULL, y = NULL) +
  theme_minimal()

p5 <- df_feat %>%
  ggplot(aes(month, selic)) +
  geom_line(color = "grey40", linewidth = 0.8) +
  labs(title = "Selic (%)", x = NULL, y = NULL) +
  theme_minimal()

p6 <- df_feat %>%
  ggplot(aes(month, ibc_br)) +
  geom_line(color = "darkorange", linewidth = 0.8) +
  labs(title = "IBC-Br", x = NULL, y = NULL) +
  theme_minimal()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 1)
```

## Correlation Matrix

```{r correlation}
cor_vars <- df_feat %>%
  select(credit_portfolio, default_rate, pix, cust, selic, ibc_br) %>%
  drop_na()

if (nrow(cor_vars) >= 3) {
  cor_mat <- cor(cor_vars)
  cat("Correlation matrix (months with all variables):\n")
  cat("N =", nrow(cor_vars), "\n\n")
  round(cor_mat, 3)
} else {
  cat("Insufficient overlapping observations for full correlation matrix.\n")

  # Compute pairwise correlations for available pairs
  cor_pairs <- df_feat %>%
    select(credit_portfolio, default_rate, pix, selic, ibc_br) %>%
    drop_na()

  if (nrow(cor_pairs) >= 3) {
    cat("Pairwise correlations (credit + macro, N =", nrow(cor_pairs), "):\n\n")
    round(cor(cor_pairs), 3)
  }
}
```

# Regression Framework

We estimate time-series OLS regressions with **Newey–West (HAC) standard errors** to address autocorrelation and heteroskedasticity. The analysis proceeds in two parts:

## Part A — Credit Growth Models

Tests whether Open Finance adoption is associated with **credit expansion**, controlling for macro conditions:

$$
g^{yoy}_{Credit,t}
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 > 0$: Open Finance adoption is associated with faster credit growth.

## Part B — Default Rate Models

Tests whether Open Finance adoption is associated with **improved credit quality**:

$$
\Delta_{12} \, Default_t
=
\beta_0
+
\beta_1 \, Selic_t
+
\beta_2 \, g^{yoy}_{IBC\text{-}Br,t}
+
\beta_3 \, \ln(\text{Customers}_t)
+
\varepsilon_t
$$

- $\beta_3 < 0$: Open Finance adoption is associated with declining default rates (better credit quality).

# Part A: Credit Growth Models

## Baseline: Credit Growth ~ Selic + IBC-Br Growth

```{r credit-baseline}
df_m_cred <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("Credit baseline sample:", nrow(df_m_cred), "months\n")
if (nrow(df_m_cred) > 0) {
  cat("Range:", as.character(min(df_m_cred$month)),
      "to", as.character(max(df_m_cred$month)), "\n")
}

if (nrow(df_m_cred) >= 5) {
  m_cred_base <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br, data = df_m_cred)
  nw_lag_cb <- max(1, floor(0.75 * nrow(df_m_cred)^(1/3)))
  nw_cred_base <- NeweyWest(m_cred_base, lag = nw_lag_cb, prewhite = FALSE)

  cat("\n--- Credit YoY Growth ~ Selic + IBC-Br Growth ---\n")
  coeftest(m_cred_base, vcov = nw_cred_base)
  cat("\nR-squared:", round(summary(m_cred_base)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for credit baseline model.\n")
}
```

## Extended: Credit Growth ~ Selic + IBC-Br + ln(Customers)

```{r credit-extended}
df_m_cred_ext <- df_feat %>%
  select(month, g_yoy_credit, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("Credit extended sample:", nrow(df_m_cred_ext), "months\n")

if (nrow(df_m_cred_ext) >= 5) {
  m_cred_ext <- lm(g_yoy_credit ~ selic + g_yoy_ibc_br + ln_cust, data = df_m_cred_ext)
  nw_lag_ce <- max(1, floor(0.75 * nrow(df_m_cred_ext)^(1/3)))
  nw_cred_ext <- NeweyWest(m_cred_ext, lag = nw_lag_ce, prewhite = FALSE)

  cat("\n--- Credit YoY Growth ~ Selic + IBC-Br Growth + ln(Customers) ---\n")
  coeftest(m_cred_ext, vcov = nw_cred_ext)
  cat("\nR-squared:", round(summary(m_cred_ext)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for credit extended model.\n")
}
```

# Part B: Default Rate Models

## Baseline: Default Rate Change ~ Selic + IBC-Br Growth

```{r default-baseline}
df_m_def <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br) %>%
  drop_na()

cat("Default rate baseline sample:", nrow(df_m_def), "months\n")

if (nrow(df_m_def) >= 5) {
  m_def_base <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br, data = df_m_def)
  nw_lag_db <- max(1, floor(0.75 * nrow(df_m_def)^(1/3)))
  nw_def_base <- NeweyWest(m_def_base, lag = nw_lag_db, prewhite = FALSE)

  cat("\n--- YoY Change in Default Rate ~ Selic + IBC-Br Growth ---\n")
  coeftest(m_def_base, vcov = nw_def_base)
  cat("\nR-squared:", round(summary(m_def_base)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for default rate baseline model.\n")
}
```

## Extended: Default Rate Change ~ Selic + IBC-Br + ln(Customers)

```{r default-extended}
df_m_def_ext <- df_feat %>%
  select(month, d_default_rate_yoy, selic, g_yoy_ibc_br, ln_cust) %>%
  drop_na()

cat("Default rate extended sample:", nrow(df_m_def_ext), "months\n")

if (nrow(df_m_def_ext) >= 5) {
  m_def_ext <- lm(d_default_rate_yoy ~ selic + g_yoy_ibc_br + ln_cust,
                  data = df_m_def_ext)
  nw_lag_de <- max(1, floor(0.75 * nrow(df_m_def_ext)^(1/3)))
  nw_def_ext <- NeweyWest(m_def_ext, lag = nw_lag_de, prewhite = FALSE)

  cat("\n--- YoY Change in Default Rate ~ Selic + IBC-Br + ln(Customers) ---\n")
  coeftest(m_def_ext, vcov = nw_def_ext)
  cat("\nR-squared:", round(summary(m_def_ext)$r.squared, 4), "\n")
} else {
  cat("Insufficient observations for default rate extended model.\n")
}
```

# Alternative Specification: Month-over-Month Growth

For robustness, we also estimate credit models using MoM growth rates.

```{r credit-mom}
# MoM Credit Growth ~ Selic + IBC-Br
df_mom_cred <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br) %>%
  drop_na()

cat("MoM credit baseline sample:", nrow(df_mom_cred), "months\n")

if (nrow(df_mom_cred) >= 5) {
  m_mom_cred <- lm(g_mom_credit ~ selic + g_mom_ibc_br, data = df_mom_cred)
  nw_mom_cred <- NeweyWest(m_mom_cred,
                           lag = max(1, floor(0.75 * nrow(df_mom_cred)^(1/3))),
                           prewhite = FALSE)

  cat("\n--- Credit MoM Growth ~ Selic + IBC-Br MoM Growth ---\n")
  coeftest(m_mom_cred, vcov = nw_mom_cred)
  cat("\nR-squared:", round(summary(m_mom_cred)$r.squared, 4), "\n")
}

# MoM Credit Growth ~ Selic + IBC-Br + ln(Customers)
df_mom_cred_ext <- df_feat %>%
  select(month, g_mom_credit, selic, g_mom_ibc_br, ln_cust) %>%
  drop_na()

cat("\nMoM credit extended sample:", nrow(df_mom_cred_ext), "months\n")

if (nrow(df_mom_cred_ext) >= 5) {
  m_mom_cred_ext <- lm(g_mom_credit ~ selic + g_mom_ibc_br + ln_cust,
                       data = df_mom_cred_ext)
  nw_mom_cred_ext <- NeweyWest(m_mom_cred_ext,
                               lag = max(1, floor(0.75 * nrow(df_mom_cred_ext)^(1/3))),
                               prewhite = FALSE)

  cat("\n--- Credit MoM Growth ~ Selic + IBC-Br MoM + ln(Customers) ---\n")
  coeftest(m_mom_cred_ext, vcov = nw_mom_cred_ext)
  cat("\nR-squared:", round(summary(m_mom_cred_ext)$r.squared, 4), "\n")
}
```

# Summary of Results

```{r summary-table}
extract_results <- function(model, nw_vcov, label) {
  ct <- coeftest(model, vcov = nw_vcov)
  tibble(
    Model = label,
    Term = rownames(ct),
    Estimate = ct[, "Estimate"],
    `Std. Error (HAC)` = ct[, "Std. Error"],
    `t value` = ct[, "t value"],
    `p value` = ct[, "Pr(>|t|)"],
    R2 = summary(model)$r.squared,
    N = nobs(model)
  )
}

# Collect all estimated models
results_list <- list()

# Credit growth models
if (exists("m_cred_base")) {
  results_list[["Credit Base (YoY)"]] <- extract_results(m_cred_base, nw_cred_base, "Credit Base (YoY)")
}
if (exists("m_cred_ext")) {
  results_list[["Credit + OF (YoY)"]] <- extract_results(m_cred_ext, nw_cred_ext, "Credit + OF (YoY)")
}
if (exists("m_mom_cred")) {
  results_list[["Credit Base (MoM)"]] <- extract_results(m_mom_cred, nw_mom_cred, "Credit Base (MoM)")
}
if (exists("m_mom_cred_ext")) {
  results_list[["Credit + OF (MoM)"]] <- extract_results(m_mom_cred_ext, nw_mom_cred_ext, "Credit + OF (MoM)")
}

# Default rate models
if (exists("m_def_base")) {
  results_list[["Default Base (YoY)"]] <- extract_results(m_def_base, nw_def_base, "Default Base (YoY)")
}
if (exists("m_def_ext")) {
  results_list[["Default + OF (YoY)"]] <- extract_results(m_def_ext, nw_def_ext, "Default + OF (YoY)")
}

results <- bind_rows(results_list)

results %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  knitr::kable(caption = "Regression Results Summary — Impact of Open Finance on Credit")
```

# Interpretation

## Credit Growth and Open Finance

The extended credit growth model tests whether Open Finance adoption ($\ln(\text{Customers})$) is associated with faster credit growth after controlling for the Selic rate and economic activity (IBC-Br). A positive and significant coefficient on $\ln(\text{Customers})$ would suggest that Open Finance is associated with credit expansion — consistent with the hypothesis that data-sharing reduces information asymmetries and enables broader lending.

Key considerations:

- **Selic**: Higher interest rates are expected to dampen credit growth ($\hat{\beta}_1 < 0$) as borrowing costs increase and both supply and demand contract.
- **IBC-Br**: Stronger economic activity should support credit growth ($\hat{\beta}_2 > 0$) through both higher demand for credit and improved creditworthiness.
- **Open Finance adoption**: A positive $\hat{\beta}_3$ is consistent with Open Finance enabling credit expansion, but could also reflect concurrent digitization trends.

## Default Rates and Open Finance

The default rate model tests whether Open Finance adoption is associated with declining default rates. A negative coefficient on $\ln(\text{Customers})$ would suggest that information-sharing improves credit quality — consistent with the hypothesis that better data allows lenders to screen borrowers more effectively.

Key considerations:

- Declining default rates during a period of Open Finance expansion could reflect improved risk assessment (causal) or simply economic recovery reducing distress (confounding).
- The national aggregate masks heterogeneity — individuals may benefit more from data portability if their credit histories were previously fragmented across institutions.

# Caveats and Limitations

1. **Short overlap period**: Open Finance customer data is available from January 2023. SCR data starts from the mid-2010s. The overlap period for models including both credit and customer data is limited.

2. **Endogeneity**: All regressors are potentially endogenous. Open Finance adoption, credit growth, and economic conditions are jointly determined. Instrumental variable or difference-in-differences approaches would be needed for causal claims.

3. **Collinearity with trend**: Customer counts are monotonically increasing and highly correlated with any time trend. Significant coefficients may capture secular trends in credit markets rather than a causal effect of Open Finance.

4. **Omitted variables**: Key omitted drivers include: regulatory changes (credit card caps, payroll lending rules), fiscal policy, housing market dynamics, commodity prices, and supply-side changes (fintech entry, bank consolidation).

5. **Aggregation**: National-level data masks heterogeneity across regions, income groups, financial institutions, and credit modalities. The impact of Open Finance may be concentrated in specific segments (e.g., low-income borrowers, consumer credit) not visible at the aggregate level.

6. **SCR data granularity**: The aggregated SCR data may group certain low-count observations under "Outros Créditos" (Other Credits), which could introduce measurement error at the modality level. National totals are less affected.

7. **Structural breaks**: The COVID-19 pandemic (2020–2021) created a major structural break in credit markets, with emergency lending programs, payment moratoria, and economic disruptions. PIX was launched during this same period, making it difficult to separate pandemic recovery effects from Open Finance effects.

# Key R Packages

| Package | Purpose |
|---------|---------|
| `tidyverse` | Data manipulation and visualization |
| `readxl` | Reading Excel files |
| `lubridate` | Date handling |
| `scales` | Axis formatting |
| `sandwich` | HAC (Newey–West) variance estimators |
| `lmtest` | Coefficient testing with robust SEs |
| `broom` | Tidy model output |
| `zoo` | Time-series utilities |
| `gridExtra` | Arranging multiple plots |

# Data Sources

| Source | Description | Location |
|--------|-------------|----------|
| SCR.Data — Carteira Ativa | Active credit portfolio (R$) | `data/carteira.csv` |
| SCR.Data — Ativo Problemático | Problem asset rate (%) | `data/ativo-problemtico.csv` |
| SCR.Data — Inadimplência | Default rate (%) | `data/inadimplncia.csv` |
| PIX / Macro / Customers | PIX volumes, Selic, IBC-Br, OF customers | `data/data_monthly.xlsx` |
